---
title: "Statistical Analysis of the Dataset"
author: "Rosario Pio Gnazzo, Antonio Pauzano"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message=FALSE, 
                      warning=FALSE, 
                      fig.align='center', fig.width = 10)
options(xts_check_TZ = FALSE)
```

```{r}
# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(corrplot)
library(gridExtra)
library(ggpubr) # For Q-Q Plot and P-P Plot

# Load the dataset
# The CSV file must be placed in the root of the repository (one level above this script).
# See data/README.md for instructions on how to obtain the dataset.
Sentiment_fr_tweet_2023 <- read_csv2("../Sentiment_fr_tweet_2023.csv")

# Always work on a copy, never on the raw dataset
data <- Sentiment_fr_tweet_2023

# Fix variable types
data <- data %>%
  mutate(userid = as.character(userid),
         username = as.character(username),
         acctdesc = as.character(acctdesc),
         location = as.character(location),
         following = as.numeric(following),
         followers = as.numeric(followers),
         totaltweets = as.numeric(totaltweets),
         tweetid = as.character(tweetid),
         retweetcount = as.numeric(retweetcount),
         favorite_count = as.numeric(favorite_count),
         text = as.character(text),
         language = as.character(language),
         favorite_count = as.numeric(favorite_count),
         
         is_retweet = as.factor(is_retweet),
         original_tweet_id = format(as.numeric(original_tweet_id), 
                                    scientific = FALSE),
         original_tweet_userid = format(as.numeric(original_tweet_userid), 
                                        scientific = FALSE),
         original_tweet_username = as.character(original_tweet_username),
         
         in_reply_to_status_id = format(as.numeric(in_reply_to_status_id), 
                                        scientific = FALSE),
         in_reply_to_user_id = format(as.numeric(in_reply_to_user_id), 
                                      scientific = FALSE),
         in_reply_to_screen_name = as.character(in_reply_to_screen_name),
         
         is_quote_status = as.factor(is_quote_status),
         quoted_status_id = format(as.numeric(quoted_status_id), 
                                   scientific = FALSE),
         quoted_status_userid = format(as.numeric(quoted_status_userid), 
                                       scientific = FALSE),
         quoted_status_username = as.character(quoted_status_username),
         
         sentiment = as.factor(sentiment),
         score = as.numeric(score)
  )
```

```{r}
# Display first rows
glimpse(data)
```

```{r}
summary(data)
```

## **1. Basic Descriptive Analysis**

-   **Central tendency measures**: mean, median, mode.

-   **Dispersion measures**: variance, standard deviation, range, interquartile range (IQR).

-   **Minimum and maximum**: to get a first idea of the value distribution.

-   **Coefficient of variation**: to assess dispersion relative to the mean.

-   **Percentiles and quartiles**: to identify data distribution.

```{r}
descr <- data %>% 
  select_if(is.numeric) %>% psych::describe()
descr %>% select(-c(vars))
```

### **General Remarks**

1.  **All variables except `score` show a strongly right-skewed distribution.**

    -   The mean is much higher than the median for most variables, indicating a long right tail.

    -   Most users have very low values, but a few extreme users push the mean upward.

2.  **Extreme outliers in `followers`, `totaltweets`, `retweetcount` and `favorite_count`**

    -   The **high kurtosis** in these variables suggests that a few individuals have values far above the mean.

3.  **`score` is the only variable with a relatively normal distribution**

    -   It has a narrow range and skewness/kurtosis values close to zero.

## **2. Preliminary Graphical Exploration**

-   **Histogram**: to observe the shape of the distribution.

-   **Q-Q (Quantile-Quantile) Plot**: to compare the empirical distribution against a theoretical one.

-   **Boxplot** + **Violin plot**: to identify the presence of outliers.

-   **Scatter plot**: to explore relationships between pairs of variables.

**Strategies adopted:**

1.  **Log-transformation** for strongly skewed variables (`followers`, `following`, `totaltweets`, `retweetcount`, `favorite_count`) to improve readability of plots.

2.  **Separate plots** for variables with and without strong skewness.

```{r}
# Log transformation for highly skewed variables
data$log_followers <- log1p(data$followers)
data$log_following <- log1p(data$following)
data$log_totaltweets <- log1p(data$totaltweets)
data$log_retweetcount <- log1p(data$retweetcount)
data$log_favorite_count <- log1p(data$favorite_count)
data$log_score <- log1p(data$score)
```

```{r}
# Detect outliers using the IQR (Tukey) method
detect_outliers_iqr <- function(x) {
  Q1 <- quantile(x, 0.25, na.rm = TRUE)
  Q3 <- quantile(x, 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  return(x < lower_bound | x > upper_bound)
}

# Detect outliers using the Z-score method
detect_outliers_zscore <- function(x) {
  z_scores <- (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
  return(abs(z_scores) > 3)
}

# Detect outliers using the Hampel method
detect_outliers_hampel <- function(x) {
  med <- median(x, na.rm = TRUE)
  mad_val <- mad(x, constant = 1.4826, na.rm = TRUE) # MAD with normalization constant
  return(abs(x - med) / mad_val > 3)
}
```

#### **Data Distribution**

Visualise the variable distributions to observe the strong skewness highlighted by the summary statistics.

```{r}
plot_density_comparison <- function(data, log_var_name, color) {
  
  # Build data for the two distributions
  df1 <- data.frame(Value = data[[log_var_name]], Type = paste(log_var_name, "(Log)"))
  df2 <- data %>% 
    filter(!detect_outliers_hampel(.data[[log_var_name]])) %>% 
    transmute(Value = .data[[log_var_name]], Type = paste(log_var_name, "(Log) - No Outliers"))
  
  # Merge datasets
  df_plot <- bind_rows(df1, df2)
  
  # Density plot
  ggplot(df_plot, aes(x = Value, color = Type)) +
    geom_histogram(aes(y = after_stat(density)),
                 bins = 30, fill = color, alpha = 0.2) +
    geom_density(linewidth=1.3) +
    scale_color_manual(values = c(paste(color,"4"), paste(color,"4"))) +
    facet_wrap(~Type, scales = "free") + 
    theme_bw() + 
    theme(legend.position = "bottom")
}
```

```{r}
p1 <- ggplot(data, aes(x = followers)) + 
  geom_histogram(aes(y = after_stat(density)),
                 bins = 30, fill = "blue", alpha = 0.5) +
  stat_function(fun = dnorm, 
                args = list(mean = mean(data$followers), 
                            sd = sd(data$followers)), color = "darkblue", size = 1) +
  theme_bw() +
  ggtitle("Followers")

p2 <- ggplot(data, aes(x = totaltweets)) + 
  geom_histogram(aes(y = after_stat(density)),
                 bins = 30, fill = "red", alpha = 0.5) +
  stat_function(fun = dnorm, 
                args = list(mean = mean(data$totaltweets), 
                            sd = sd(data$totaltweets)), color = "darkred", size = 1) +
  theme_bw() +
  ggtitle("Total Tweets")

p3 <- ggplot(data, aes(x = following)) + 
  geom_histogram(aes(y = after_stat(density)),
                 bins = 30, fill = "purple", alpha = 0.5) +
  stat_function(fun = dnorm, 
                args = list(mean = mean(data$following), 
                            sd = sd(data$following)), color = "purple4", size = 1) +
  theme_bw() +
  ggtitle("Following")

p4 <- ggplot(data, aes(x = retweetcount)) + 
  geom_histogram(aes(y = after_stat(density)),
                 bins = 30, fill = "brown", alpha = 0.5) +
  stat_function(fun = dnorm, 
                args = list(mean = mean(data$retweetcount), 
                            sd = sd(data$retweetcount)), color = "brown4", size = 1) +
  theme_bw() +
  ggtitle("Retweetcount")

p5 <- ggplot(data, aes(x = favorite_count)) + 
  geom_histogram(aes(y = after_stat(density)),
                 bins = 30, fill = "orange", alpha = 0.5) +
  stat_function(fun = dnorm, 
                args = list(mean = mean(data$favorite_count), 
                            sd = sd(data$favorite_count)), color = "orange4", size = 1) +
  theme_bw() +
  ggtitle("Favouritecount")

p6 <- ggplot(data, aes(x = score)) + 
  geom_histogram(aes(y = after_stat(density)),
                 bins = 30, fill = "green", alpha = 0.5) +
  stat_function(fun = dnorm, 
                args = list(mean = mean(data$score), 
                            sd = sd(data$score)), color = "darkgreen", size = 1) +
  theme_bw() +
  ggtitle("Score")

grid.arrange(p1, p2, p3, p4, p5, p6, ncol=3)
```

As already shown by the summary statistics, most variable distributions are strongly right-skewed and contain anomalous values. We therefore apply **log transformations** and **outlier detection** to obtain clearer visualisations.

#### Log Transformation

We observe how the distribution and outliers change by visualising the boxplot for each variable on a logarithmic scale.

```{r}
# Reshape data to long format
df_long <- data %>%
  select(log_followers, log_totaltweets, log_following, 
         log_retweetcount, log_favorite_count, log_score) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value")

color_palette <- c(
  "log_followers" = "blue",
  "log_totaltweets" = "red",
  "log_following" = "purple",
  "log_retweetcount" = "brown",
  "log_favorite_count" = "orange",
  "log_score" = "green"
)

# Boxplot
  ggplot(df_long, aes(y = Variable, x = Value, fill = Variable)) +
  geom_boxplot(alpha = 0.5, outlier.alpha = 0.5, notch = TRUE) +
  theme_bw() +
  ggtitle("Variable Boxplots") +
  scale_fill_manual(values = color_palette)
```

#### **Outlier Identification**

Three methods are applied to detect outliers:

-   **Tukey IQR rule**: values beyond 1.5*IQR from the quartiles.

-   **Z-score**: values beyond \u00b13 standard deviations.

-   **Hampel method**: based on the median and median absolute deviation (MAD).

```{r}
# Apply outlier detection to all numeric variables
outliers_iqr <- data %>% mutate(across(where(is.numeric), detect_outliers_iqr))
outliers_zscore <- data %>% mutate(across(where(is.numeric), detect_outliers_zscore))
outliers_hampel <- data %>% mutate(across(where(is.numeric), detect_outliers_hampel))

# Summary table of outlier counts per variable and method
summary_outliers <- data.frame(
  Variable = names(data %>% select(where(is.numeric))),
  IQR_Outliers = sapply(data %>% select(where(is.numeric)), function(x) sum(detect_outliers_iqr(x), na.rm = TRUE)),
  ZScore_Outliers = sapply(data %>% select(where(is.numeric)), function(x) sum(detect_outliers_zscore(x), na.rm = TRUE)),
  Hampel_Outliers = sapply(data %>% select(where(is.numeric)), function(x) sum(detect_outliers_hampel(x), na.rm = TRUE))
)

print(summary_outliers)
```

-   **Variabili non trasformate**

    -   **`following`, `followers`, `totaltweets`, `retweetcount`, `favorite_count`** hanno un numero elevato di outlier secondo tutti i metodi.

    -   In particolare, **`retweetcount`** e **`favorite_count`** hanno una forte asimmetria, come si deduce dall’elevato numero di outlier Z-score e Hampel.

    -   **`score`** non presenta outlier in nessun metodo, suggerendo una distribuzione più regolare.

-   **Effetto della trasformazione logaritmica**

    -   Dopo la trasformazione logaritmica, il numero di outlier si riduce drasticamente per **tutte** le variabili, specialmente per `log_followers`, `log_following`, `log_totaltweets`.

    -   **`log_retweetcount`** ha **zero outlier** con ogni metodo, suggerendo che la trasformazione ha eliminato gran parte dell'estrema variabilità nei dati originali.

    -   Questo suggerisce che i dati originali erano molto sbilanciati e altamente asimmetrici, mentre la trasformazione logaritmica ha ridotto questa distorsione.

```{r}
library(reshape2)

melted_outliers <- melt(summary_outliers, id.vars = "Variable")
ggplot(melted_outliers, aes(x = Variable, y = value, fill = variable)) + 
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Outlier count by method and variable")
```

Distributions after removing outliers (Hampel method):

```{r, fig.height=15, fig.width=25}
p1<-plot_density_comparison(data,"log_followers", "blue")
p2<-plot_density_comparison(data,"log_totaltweets", "red")
p3<-plot_density_comparison(data,"log_following", "purple")
p4<-plot_density_comparison(data,"log_retweetcount", "brown")
p5<-plot_density_comparison(data, "log_score", "green")

grid.arrange(p1, p2, p3, p4, p5, ncol=2)
```

```{r, fig.height=10, fig.width=10}
df <- data %>% filter(!detect_outliers_hampel(log_followers))
p7<-ggqqplot(df$log_followers, 
         color = "blue",
         ggtheme = theme_bw(),
         title = "Q-Q plot: log_followers & No-Outliers")

df <- data %>% filter(!detect_outliers_hampel(log_totaltweets))
p8 <- ggqqplot(df$log_totaltweets, 
         color = "red",
         ggtheme = theme_bw(),
         title = "Q-Q plot: log_totaltweets & No-Outliers")

df <- data %>% filter(!detect_outliers_hampel(log_following)) 
p9 <- ggqqplot(df$log_following, 
         color = "purple",
         ggtheme = theme_bw(),
         title = "Q-Q plot: log_following & No-Outliers")

df <- data %>% filter(!detect_outliers_hampel(log_retweetcount))
p10 <- ggqqplot(df$log_retweetcount, 
         color = "brown",
         ggtheme = theme_bw(),
         title = "Q-Q plot: log_retweetcount & No-Outliers")

df <- data %>% filter(!detect_outliers_hampel(log_score))
p12 <- ggqqplot(df$log_score, 
         color = "green",
         ggtheme = theme_bw(),
         title = "Q-Q plot: log_score & No-Outliers")

grid.arrange(p7,
             p8,
             p9,
             p10, 
             p12, ncol = 3)
```

**Chi-Square Normality Test:**

```{r}
## Manual chi-square test for normality on log_followers
x <- data %>% filter(!detect_outliers_hampel(log_followers)) %>%
  select(log_followers) %>%
  summarise( N = n(),
             M = mean(log_followers),
             sd = sd(log_followers))

a <- numeric(4)
for(i in 1:4){
  a[i] <- qnorm(0.2*i, mean = x$M, sd=x$sd)
}

a

# Count observations falling in each interval
r <- 5
nint <- numeric(r)
nint[1] <- data %>% 
  filter(!detect_outliers_hampel(log_followers)) %>%
  select(log_followers) %>%
  filter(log_followers < a[1]) %>%
  summarise( N = n())%>%
  pull(N)
nint[2] <- data %>% 
  filter(!detect_outliers_hampel(log_followers)) %>%
  select(log_followers) %>%
  filter(log_followers >= a[1] & log_followers < a[2]) %>%
  summarise( N = n())%>%
  pull(N)
nint[3] <- data %>% 
  filter(!detect_outliers_hampel(log_followers)) %>%
  select(log_followers) %>%
  filter(log_followers >= a[2] & log_followers < a[3]) %>%
  summarise( N = n())%>%
  pull(N)
nint[4] <- data %>% 
  filter(!detect_outliers_hampel(log_followers)) %>%
  select(log_followers) %>%
  filter(log_followers >= a[3] & log_followers < a[4]) %>%
  summarise( N = n())%>%
  pull(N)
nint[5] <- data %>% 
  filter(!detect_outliers_hampel(log_followers)) %>%
  select(log_followers) %>%
  filter(log_followers >= a[4]) %>%
  summarise( N = n())%>%
  pull(N)


chisq.test(nint, p = rep(0.2, length(nint)))

df <- data %>% 
  filter(!detect_outliers_hampel(log_followers)) %>%
  select(log_followers)

nortest::lillie.test(df$log_followers)

data %>% select(log_followers, log_following, log_totaltweets, log_retweetcount, log_favorite_count, log_score)
```

```{r}
remove_extreme_values <- function(data, lower_quantile = 0.05, upper_quantile = 0.95) {
  q_lower <- quantile(data, lower_quantile, na.rm = TRUE)
  q_upper <- quantile(data, upper_quantile, na.rm = TRUE)
  
  # Filter values between specified quantiles
  filtered_data <- data[data >= q_lower & data <= q_upper]
  
  return(filtered_data)
}
```

#### **Scatter Plot: Relationships between Variables**

Check for linear correlations between numerical variables.

```{r, fig.height=10}
library(GGally)

data %>%
  select(log_followers, log_following, log_totaltweets, log_retweetcount, log_favorite_count, log_score) %>%
  GGally::ggpairs(
    upper = list(continuous = wrap("cor", size = 4)),
    lower = list(continuous = wrap("smooth", color = "gray60")),
    diag = list(continuous = wrap("densityDiag", fill = "lightblue"))
  ) +
  theme_bw()
```

Linear relationships are visible among several numerical variables. Let us inspect them more closely:

```{r}
# Select numeric variables
data_selected <- data %>% select(followers,following, totaltweets, retweetcount, favorite_count, score)

# Pearson correlation matrix
cor_pearson <- cor(data_selected, use = "complete.obs", method = "pearson")

# Correlation heatmap (Pearson)
corrplot(cor_pearson, method = "color", type = "lower", order = "hclust", 
         col = colorRampPalette(c("blue3", "white", "red3"))(200), 
         diag = FALSE,
         tl.cex = 0.8, tl.col = "black",
         addCoef.col = "black",
         number.cex = 0.7)
```

Key observations:

**Strong correlations**

-   **`log_followers` and `log_following` (0.69)** - Users with more followers tend to follow more accounts.

-   **`log_followers` and `log_totaltweets` (0.65)** - Users with more followers tend to have posted more tweets.

**Moderate correlations**

-   **`log_totaltweets` and `log_following` (0.47)** - Users who follow many accounts also tend to tweet more.

-   **`log_followers` and `log_favorite_count` (0.27)** - A slight positive correlation: more followers may lead to more likes.

**Weak or negative correlations**

-   **`log_retweetcount` and `log_followers` (-0.16)** - Weak negative: more followers do not necessarily mean more retweets.

-   **`log_retweetcount` and `log_favorite_count` (-0.20)** - Inverse tendency between retweets and likes.

-   **`score` shows very low correlations** with all other variables, suggesting it does not depend linearly on the engagement metrics.

#### Linear Regression Model (followers)

```{r}
model <- lm(log_followers ~ log_following + log_totaltweets + log_retweetcount + log_favorite_count + log_score, data = data)

# Model summary
summary(model)

forecast::checkresiduals(model)
```

```{r}
model <- lm(log_followers ~ log_following + log_totaltweets + log_retweetcount + log_favorite_count + log_score + sentiment + is_retweet + is_quote_status, data = data)

# Model summary
summary(model)
```

#### Multiple Linear Regression (Score)

```{r}
model1 <- lm(score ~ log_followers + log_following + log_totaltweets + 
              log_retweetcount + log_favorite_count, data = data)

# Model summary
summary(model1)
```

```{r}
df <- data %>%
  select(score, log_followers, log_following, log_totaltweets, log_retweetcount, log_favorite_count, is_retweet, is_quote_status, sentiment)

model2 <- lm(score ~ ., data = df)

# Model summary
summary(model2)
```

**Model 1 (log variables only)**

-   **`log_retweetcount` (+0.0129)** - The retweet count has the largest positive impact on `score`.

-   **`log_favorite_count` (+0.0046)** - Likes also have a positive effect.

-   **`log_followers` (-0.0033)** - Counter-intuitively, more followers is associated with a lower score.

-   **Low predictive power** (`R-sq Adjusted = 0.0247`), suggesting other variables not included are more important.

**Model 2 (log variables + categorical)**

-   **`is_retweetTRUE` (-0.0483)** - Being a retweet significantly decreases `score`.

-   **`is_quote_statusTRUE` (+0.0203)** - Being a quote tweet increases `score`.

-   **`sentimentneu` (-0.2141)** and **`sentimentpos` (-0.1426)** - Neutral and positive tweets have a **negative** impact on `score` relative to negative tweets (baseline), suggesting negative tweets may drive more engagement.

-   **`log_retweetcount` (+0.0142)** - Again, retweet count is the strongest positive predictor.

-   **`log_favorite_count` (-0.0101)** - Unlike in Model 1, this predictor now shows a negative effect.

Model 2 is substantially better than Model 1:

-   **Higher R-sq Adjusted** (0.2499 vs 0.0247): explains a larger proportion of score variability.

-   **Lower Residual Standard Error** (0.1571 vs 0.1791): more accurate predictions.

-   **More significant variables**, including the categorical predictors `is_retweet`, `is_quote_status` and the sentiment levels.

**Conclusions**

1.  **Model 2 is clearly superior** in terms of explanatory power and accuracy.

2.  **Categorical variables (retweet, quote status, sentiment) are strongly influential** and would be missed in Model 1.

3.  **Negative sentiment appears to favour a higher score** — an interesting finding for engagement analysis.

Without log transformation, the predictive power of the model collapses:

```{r}
df <- data %>%
  select(following, followers, totaltweets, retweetcount, favorite_count, is_retweet, is_quote_status, sentiment, score)

model <- lm(score ~ ., data = df)

summary(model)
```

#### Multiple Logistic Regression (Sentiment)

**General Interpretation**

-   The model uses **negative sentiment as the reference category**.

-   The two rows in the coefficients table represent:

    -   **neu** - log-odds of a tweet being **neutral** vs negative.

    -   **pos** - log-odds of a tweet being **positive** vs negative.

-   Positive coefficients increase the probability of belonging to that category relative to negative; negative coefficients indicate higher probability of being negative.

```{r}
library(nnet)

df <- data %>%
  select(log_following, log_followers, log_totaltweets, log_retweetcount, log_favorite_count, is_retweet, is_quote_status, sentiment, log_score)

# Multinomial model: sentiment as dependent variable
model_multinom <- multinom(sentiment ~ ., data = df)

# Model summary
summary(model_multinom)
```

```{r}
# Summary table of coefficient direction per sentiment class
tabella_confronto <- data.frame(
  Factor = c("High score", "More followers", "More following", "More total tweets",
              "More retweets received", "More likes received", "Is a retweet", "Is a quote"),
  Negative = c("Increases", "Decreases", "Decreases", "Increases",
               "Increases", "Increases", "Decreases", "Increases"),
  Neutral  = c("Decreases", "Increases", "Decreases", "Decreases",
               "Decreases", "Decreases", "Decreases", "Decreases"),
  Positive = c("Decreases", "Increases", "Decreases", "Decreases",
               "Decreases", "Decreases", "Decreases", "Increases")
)

tabella_confronto
```

**1. Negative Tweets (Reference Category)**

Factors that increase the probability of negative sentiment:

-   **Higher score** - The higher the score, the more likely the tweet is negative.

-   **More total tweets by the user** - Very active users are more likely to write negative tweets.

-   **Fewer likes received** - Less appreciated tweets tend to be more negative.

-   **Neutral and positive sentiments consistently show negative coefficients relative to negative**, suggesting that **negative is the default state in the absence of strong external influences**.

Possible explanation: negative tweets tend to be more polarising and shareable (higher engagement); highly active users may be more likely to express criticism.

**2. Neutral Tweets**

Factors that increase the probability of neutral vs negative:

-   **More followers** - Users with more followers tend to write neutral tweets.

-   **Fewer retweets received** - Tweets with fewer shares are more often neutral.

-   **Not a retweet or quote** - Original tweets tend to be more neutral.

Factors that decrease the probability of neutral:

-   **High score** - A high score reduces the probability of neutrality.

-   **More likes** - More appreciated tweets tend to be less neutral (more polarised).

-   **Quote tweets** - Quote status reduces the chance of neutral classification.

Possible explanation: users with large audiences may aim for safer, neutral communication; original tweets are less polarised than retweets or quotes.

**3. Positive Tweets**

Factors that increase the probability of positive vs negative:

-   **More followers** - Users with more followers are more likely to post positive tweets.

-   **Quote tweets** - Quoting another tweet correlates with positive sentiment.

-   **Slightly lower score** - Positive tweets tend to have a slightly lower score than negative ones.

Factors that decrease the probability of positive:

-   **More total tweets** - Very prolific tweeters are less likely to be positive.

-   **More likes** - Unexpectedly, tweets with more likes are **less likely to be positive** relative to negative.

-   **Retweets** - Being a retweet reduces the chance of positive classification.

Possible explanation: users with large audiences may have more encouraging communities; quote tweets are often used to add positive commentary; retweets are frequently used to amplify criticism.

### **3. Statistical Inference**

-   Comparison with theoretical distributions (normal, exponential, gamma, beta, etc.).

-   **Understanding the data distribution** allows us to:

    -   Grasp the shape (normal, exponential, gamma, beta, etc.).

    -   Simplify interpretation of statistical properties (skewness, variance, etc.).

    **Example:** If a variable follows a **normal** distribution, classical parametric tests apply directly.

-   **Selection of appropriate statistical tests**

    -   Some tests **require distributional assumptions** (e.g., normality for the t-test or ANOVA).

    -   When normality cannot be assumed, **non-parametric tests** are more appropriate (e.g., Wilcoxon, Mann-Whitney).

    **Example:** When the distribution is highly skewed, a **Wilcoxon test** is more suitable than a t-test.

    -   **Hypothesis tests**:

        -   **t-test**: compare a variable mean to a known value, or compare two groups.

        -   **ANOVA**: compare means across more than two groups.

        -   **Chi-square**: test independence between two categorical variables.

    -   **Confidence intervals**:

        -   Construct confidence intervals for means or proportions to quantify estimation uncertainty.

```{r}
df <- data %>% select(log_followers, log_following, log_totaltweets, log_retweetcount, log_favorite_count, log_score)

# 1. Wilcoxon test for the median
# Reference value: sample median
median_values <- df %>% summarise(across(everything(), median, na.rm = TRUE))

# Apply Wilcoxon test to each variable
wilcoxon_results <- lapply(names(df), function(var) {
  test <- wilcox.test(df[[var]], mu = median_values[[var]], alternative = "two.sided")
  data.frame(Variable = var, P_value = test$p.value, Estimated_Median = median_values[[var]])
})
wilcoxon_results <- bind_rows(wilcoxon_results)
print(wilcoxon_results)
```

```{r}
# 2. Confidence intervals for the median: classic and quantile methods
confidence_intervals <- lapply(names(df), function(var) {
  # Classic method (assuming normality)
  n <- sum(!is.na(df[[var]]))
  med <- median(df[[var]], na.rm = TRUE)
  se <- 1.253 * sd(df[[var]], na.rm = TRUE) / sqrt(n)
  ic_classic <- c(med - 1.96 * se, med + 1.96 * se)
  
  # Quantile-based method
  ic_quantile <- quantile(df[[var]], probs = c(0.025, 0.975), na.rm = TRUE)
  
  data.frame(Variable = var,
             IC_Classic_Lower = ic_classic[1], IC_Classic_Upper = ic_classic[2],
             IC_Quantile_Lower = ic_quantile[1], IC_Quantile_Upper = ic_quantile[2])
})
confidence_intervals <- bind_rows(confidence_intervals)
print(confidence_intervals)

```

### Categorical Analysis

```{r}
# Function to analyse a categorical variable
analysis_categorical <- function(df, column) {
  table_out <- df %>%
    count(!!sym(column)) %>%
    mutate(percentage = (n / sum(n)) * 100) %>%
    arrange(desc(n))
  
  n_unique <- n_distinct(df[[column]])
  mode_val <- table_out[[column]][which.max(table_out$n)]
  mode_pct <- max(table_out$percentage)
  
  table_out <- table_out %>%
    mutate(Variable = column, 
           Unique_Categories = n_unique,
           Mode = mode_val, 
           Mode_Percentage = mode_pct) %>%
    select(Variable, everything())
  
  return(table_out)
}

# Select only categorical columns (factors)
categorical_vars <- names(data)[sapply(data, is.factor)]

# Apply function to all categorical variables
results_categorical <- bind_rows(lapply(categorical_vars, function(col) analysis_categorical(data, col)))

print(results_categorical)
```

```{r}
library(lubridate)
dataExtract <- data %>%
  filter(!is.na(extractedts)) %>%
  mutate(
    hour_of_day = hour(extractedts),
    day_of_week = wday(extractedts, label = TRUE, abbr = FALSE),
    interval_sec = as.numeric(difftime(extractedts, lag(extractedts), units = "secs"))
  )

# Frequency tables and mean extraction interval
print(table(dataExtract$hour_of_day)/length(dataExtract$hour_of_day))
print(table(dataExtract$day_of_week)/length(dataExtract$day_of_week))
print(mean(na.omit(dataExtract$interval_sec), na.rm = TRUE))
```
