---
title: "Granular Temporal Analysis - Tweet Count"
author: "Rosario Pio Gnazzo, Antonio Pauzano"
date: "2025-01-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message=FALSE, 
                      warning=FALSE, 
                      fig.align='center', fig.width = 10)
options(xts_check_TZ = FALSE)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(highfrequency)
library(xts)
library(forecast)
library(lubridate)
library(gridExtra)
library(urca)
library(tibble)
```

```{r}
# Load the dataset.
# The CSV file must be placed in the root of the repository (one level above this script).
# See data/README.md for instructions on how to obtain the dataset.
Sentiment_fr_tweet_2023 <- read_csv2("../Sentiment_fr_tweet_2023.csv")

dataset <- Sentiment_fr_tweet_2023

dataset <- dataset %>%
  mutate(userid = as.character(userid),
         following = as.numeric(following),
         followers = as.numeric(followers),
         totaltweets = as.numeric(totaltweets),
         tweetid = as.character(tweetid),
         retweetcount = as.numeric(retweetcount),
         favorite_count = as.numeric(favorite_count),
         is_retweet = as.factor(ifelse(is_retweet == FALSE, 0, 1)),
         is_quote_status = as.factor(ifelse(is_quote_status == FALSE, 0, 1)),
         sentiment = as.factor(sentiment),
         score = as.numeric(score),
         score_std = scale(score))
```

# Temporal Analysis of Tweet Volume

Un idea potrebbe essere quella di osservare la distribuzione nel tempo dei post su Twitter contando il numero di occorrenze per sentiment in un dato sotto-intervallo di tempo.

```{r, include=FALSE}
df_occ10 <- dataset %>%   
  mutate(interval = floor_date(tweetcreatedts, "10 minutes")) %>% 
  group_by(interval, sentiment) %>% 
  summarise(count = n(), .groups = "drop") %>%
  arrange(interval, sentiment)

df_occ30 <- dataset %>%
  mutate(interval = floor_date(tweetcreatedts, "30 minutes")) %>% 
  group_by(interval, sentiment) %>% 
  summarise(count = n(), .groups = "drop") %>%
  arrange(interval, sentiment)

df_occ1h <- dataset %>%
  mutate(interval = floor_date(tweetcreatedts, "1 hours")) %>% 
  group_by(interval, sentiment) %>% 
  summarise(count = n(), .groups = "drop") %>%
  arrange(interval, sentiment)

df_occ6h <- dataset %>%
  mutate(interval = floor_date(tweetcreatedts, "6 hours")) %>% 
  group_by(interval, sentiment) %>% 
  summarise(count = n(), .groups = "drop") %>%
  arrange(interval, sentiment)
```

```{r, fig.width=8, fig.height=8}
# Grafico a linee 
p1 <- ggplot(df_occ10, aes(x = interval, y = count, color = sentiment)) +   
  geom_line() +   
  labs( title = "Granularity: 10 minutes", 
        x = "Intervalli di 10 Minuti", 
        y = "Number of Tweets", color = "Sentiment") +   
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_datetime(date_breaks = "6 hours", date_labels = "%d-%b %H:%M")

p2 <- ggplot(df_occ30, aes(x = interval, y = count, color = sentiment)) +   
  geom_line() +   
  labs( title = "Granularity: 30 minutes", 
        x = "Intervalli di 30 Minuti", 
        y = "Number of Tweets", color = "Sentiment") +   
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_datetime(date_breaks = "6 hours", date_labels = "%d-%b %H:%M")

p3 <- ggplot(df_occ1h, aes(x = interval, y = count, color = sentiment)) +   
  geom_line() +   
  labs( title = "Granularity: 1 hour", 
        x = "Intervalli di 1 ora", 
        y = "Number of Tweets", color = "Sentiment") +   
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_datetime(date_breaks = "6 hours", date_labels = "%d-%b %H:%M")

grid.arrange(p1, p2, p3, ncol=1)
```

The plot clearly captures the presence of a temporal periodic structure in the data, which persists across all levels of analysis. Let us examine this structure more closely.

## Periodicity Analysis of the Time Series

```{r, include=FALSE}
# Reshape to wide format for convenience 
df_occ_wide10 <- df_occ10 %>%   
  pivot_wider(names_from = sentiment, values_from = count, values_fill = 0)  
df_occ_wide30 <- df_occ30 %>%   
  pivot_wider(names_from = sentiment, values_from = count, values_fill = 0) 
df_occ_wide1h <- df_occ1h %>%   
  pivot_wider(names_from = sentiment, values_from = count, values_fill = 0) 

# Create time series for un sentiment specifico (es. "neg") 
neg_ts10 <- ts(df_occ_wide10$neg, start = c(2023, 1), frequency = 144) # 144 intervalli di 10 minuti al giorno
neg_ts30 <- ts(df_occ_wide30$neg, start = c(2023, 1), frequency = 48) # 48 intervalli di 30 minuti al giorno 
neg_ts1h <- ts(df_occ_wide1h$neg, start = c(2023, 1), frequency = 24) # 24 intervalli di 1 pra al giorno

# Create time series for un sentiment specifico (es. "neu") 
neu_ts10 <- ts(df_occ_wide10$neu, start = c(2023, 1), frequency = 144) # 144 intervalli di 10 minuti al giorno
neu_ts30 <- ts(df_occ_wide30$neu, start = c(2023, 1), frequency = 48) # 48 intervalli di 30 minuti al giorno 
neu_ts1h <- ts(df_occ_wide1h$neu, start = c(2023, 1), frequency = 24) # 24 intervalli di 1 pra al giorno

# Create time series for un sentiment specifico (es. "pos") 
pos_ts10 <- ts(df_occ_wide10$pos, start = c(2023, 1), frequency = 144) # 144 intervalli di 10 minuti al giorno
pos_ts30 <- ts(df_occ_wide30$pos, start = c(2023, 1), frequency = 48) # 48 intervalli di 30 minuti al giorno 
pos_ts1h <- ts(df_occ_wide1h$pos, start = c(2023, 1), frequency = 24) # 24 intervalli di 1 pra al giorno
```

### Negative Tweets

```{r, include=FALSE}
p1<-ggseasonplot(neg_ts10)+
  scale_color_manual(
    values = c("red", "blue", "green", "purple", "orange", "cyan", "yellow"), # Colori personalizzati
    labels = c("day1", "day2", "day3", "day4", "day5", "day6", "day7"), # Nuove etichette
    name = "Days")+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  xlab("Minutes")+
  ggtitle("")+
  theme_light()+
  theme(legend.position = "top")

p2<-ggseasonplot(neg_ts30)+
  scale_color_manual(
    values = c("red", "blue", "green", "purple", "orange", "cyan", "yellow"), # Colori personalizzati
    labels = c("day1", "day2", "day3", "day4", "day5", "day6", "day7"), # Nuove etichette
    name = "Days")+
  xlab("Minutes")+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  ggtitle("")+
  theme_light()+
  theme(legend.position = "top")

p3<-ggseasonplot(neg_ts1h)+
  scale_color_manual(
    values = c("red", "blue", "green", "purple", "orange", "cyan", "yellow"), # Colori personalizzati
    labels = c("day1", "day2", "day3", "day4", "day5", "day6", "day7"), # Nuove etichette
    name = "Days" # Nuovo titolo della leggenda
  )+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()+ggtitle("")+
  theme(legend.position = "top")

p4 <- ggAcf(neg_ts10, lag.max = 300)+
  ggtitle("ACF - 10 min") + theme_light()

p5 <- ggAcf(neg_ts30, lag.max = 100)+
  ggtitle("ACF - 30 min") + theme_light()

p6<-ggAcf(neg_ts1h, lag.max = 50)+
  ggtitle("ACF - 1 ora") + theme_light()
```

#### **10-minute granularity:**

```{r, fig.height=5, fig.width=10}
grid.arrange(p1, p4, ncol=2)
Box.test(neg_ts10, type = "Ljung-Box")
```

#### **30-minute granularity:**

```{r, fig.height=5, fig.width=10}
grid.arrange(p2, p5, ncol=2)
Box.test(neg_ts30, type = "Ljung-Box") 
```

#### **1-hour granularity:**

```{r, fig.height=5, fig.width=10}
grid.arrange(p3, p6, ncol=2)
Box.test(neg_ts1h, type = "Ljung-Box")
```

### Neutral Tweets

```{r, include=FALSE}
p1<-ggseasonplot(neu_ts10)+
  scale_color_manual(
    values = c("red", "blue", "green", "purple", "orange", "cyan", "yellow"), # Colori personalizzati
    labels = c("day1", "day2", "day3", "day4", "day5", "day6", "day7"), # Nuove etichette
    name = "Days")+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  xlab("Minutes")+
  ggtitle("")+
  theme_light()+
  theme(legend.position = "top")

p2<-ggseasonplot(neu_ts30)+
  scale_color_manual(
    values = c("red", "blue", "green", "purple", "orange", "cyan", "yellow"), # Colori personalizzati
    labels = c("day1", "day2", "day3", "day4", "day5", "day6", "day7"), # Nuove etichette
    name = "Days")+
  xlab("Minutes")+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  ggtitle("")+
  theme_light()+
  theme(legend.position = "top")

p3<-ggseasonplot(neu_ts1h)+
  scale_color_manual(
    values = c("red", "blue", "green", "purple", "orange", "cyan", "yellow"), # Colori personalizzati
    labels = c("day1", "day2", "day3", "day4", "day5", "day6", "day7"), # Nuove etichette
    name = "Days" # Nuovo titolo della leggenda
  )+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()+ggtitle("")+
  theme(legend.position = "top")

p4 <- ggAcf(neu_ts10, lag.max = 300)+
  ggtitle("ACF") + theme_light()

p5 <- ggAcf(neu_ts30, lag.max = 100)+
  ggtitle("ACF") + theme_light()

p6<-ggAcf(neu_ts1h, lag.max = 50)+
  ggtitle("ACF") + theme_light()
```

#### **10-minute granularity:**

```{r, fig.height=3}
grid.arrange(p1, p4, ncol=2)
Box.test(neu_ts10, type = "Ljung-Box")
```

#### **30-minute granularity:**

```{r, fig.height=3}
grid.arrange(p2, p5, ncol=2)
Box.test(neu_ts10, type = "Ljung-Box")
```

#### **1-hour granularity:**

```{r, fig.height=3}
grid.arrange(p3, p6, ncol=2)
Box.test(neu_ts10, type = "Ljung-Box")
```

### Positive Tweets

```{r, include=FALSE}
p1<-ggseasonplot(pos_ts10)+
  scale_color_manual(
    values = c("red", "blue", "green", "purple", "orange", "cyan", "yellow"), # Colori personalizzati
    labels = c("day1", "day2", "day3", "day4", "day5", "day6", "day7"), # Nuove etichette
    name = "Days")+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  xlab("Minutes")+
  ggtitle("")+
  theme_light()+
  theme(legend.position = "top")

p2<-ggseasonplot(pos_ts30)+
  scale_color_manual(
    values = c("red", "blue", "green", "purple", "orange", "cyan", "yellow"), # Colori personalizzati
    labels = c("day1", "day2", "day3", "day4", "day5", "day6", "day7"), # Nuove etichette
    name = "Days")+
  xlab("Minutes")+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  ggtitle("")+
  theme_light()+
  theme(legend.position = "top")

p3<-ggseasonplot(pos_ts1h)+
  scale_color_manual(
    values = c("red", "blue", "green", "purple", "orange", "cyan", "yellow"), # Colori personalizzati
    labels = c("day1", "day2", "day3", "day4", "day5", "day6", "day7"), # Nuove etichette
    name = "Days" # Nuovo titolo della leggenda
  )+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()+ggtitle("")+
  theme(legend.position = "top")

p4 <- ggAcf(pos_ts10, lag.max = 300)+
  ggtitle("ACF") + theme_light()

p5 <- ggAcf(pos_ts30, lag.max = 100)+
  ggtitle("ACF") + theme_light()

p6<-ggAcf(pos_ts1h, lag.max = 50)+
  ggtitle("ACF") + theme_light()
```

#### **10-minute granularity:**

```{r, fig.height=3}
grid.arrange(p1, p4, ncol=2) 
Box.test(pos_ts10, type = "Ljung-Box")
```

#### **30-minute granularity:**

```{r, fig.height=3}
grid.arrange(p2, p5, ncol=2) 
Box.test(pos_ts10, type = "Ljung-Box")
```

#### **1-hour granularity:**

```{r, fig.height=3}
grid.arrange(p3, p6, ncol=2) 
Box.test(pos_ts10, type = "Ljung-Box")
```

A clear periodic structure is visible in the data. Furthermore, the Ljung-Box test, which assesses whether the correlations detected at various lags are statistically significant, is always significant (p-value \< 0.05), confirming the presence of significant autocorrelations driven by the underlying periodic structure.

## Statistical Models (Negative Series Only)

We search for the best statistical model to describe the time series.

### Decomposition Models

Decomposition models separate a time series into its structural components: *trend*, *seasonality*, and *remainder* (residuals). This allows analysis of each component independently.

#### STL Decomposition

STL (Seasonal and Trend decomposition using Loess) is a robust, flexible decomposition method that handles any seasonality type and is resistant to outliers.

```{r}
# Time series decomposition 
# Observe the seasonal component to detect repeating patterns  
neg_ts10 %>% 
  stl(t.window = 1008, s.window = 144, robust = TRUE) -> fit10

neg_ts30 %>% 
  stl(t.window = 336, s.window = 48, robust = TRUE) -> fit30

neg_ts1h %>% 
  stl(t.window = 168, s.window = 24, robust = TRUE) -> fit1h

p1<-autoplot(neg_ts10, series="Data") +   
  autolayer(trendcycle(fit10), series="Trend") +   
  autolayer(seasadj(fit10), series="Seasonally Adjusted") +   
  xlab("Time") + 
  ylab("Number of tweets") +   
  ggtitle("STL Decomposition: negative tweets every 10 minutes") + 
  scale_colour_manual(values=c("gray","blue","red"),breaks=c("Data","Seasonally Adjusted","Trend"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

p2<-autoplot(neg_ts30, series="Data") +   
  autolayer(trendcycle(fit30), series="Trend") +   
  autolayer(seasadj(fit30), series="Seasonally Adjusted") +   
  xlab("Time") + 
  ylab("Number of tweets") +   
  ggtitle("STL Decomposition: negative tweets every 30 minutes") + 
  scale_colour_manual(values=c("gray","blue","red"),breaks=c("Data","Seasonally Adjusted","Trend"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

p3<-autoplot(neg_ts1h, series="Data") +   
  autolayer(trendcycle(fit1h), series="Trend") +   
  autolayer(seasadj(fit1h), series="Seasonally Adjusted") +   
  xlab("Time") + 
  ylab("Number of tweets") +   
  ggtitle("STL Decomposition: negative tweets every hour") +   
  scale_colour_manual(values=c("gray","blue","red"),breaks=c("Data","Seasonally Adjusted","Trend"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

grid.arrange(p1, p2, p3, ncol=1)
```

The STL decomposition clearly reveals the periodic pattern in the negative tweet series. The trend component shows a gradual variation over the observation period, while the seasonal component captures the recurring daily pattern.

#### Classical Additive Decomposition

The additive decomposition model assumes that the components combine additively: $y_t = T_t + S_t + R_t$, where trend, seasonality, and remainder sum to the observed value.

```{r}
neg_ts10 %>% decompose(type="additive") -> fit10A 
neg_ts30 %>% decompose(type="additive") -> fit30A 
neg_ts1h %>% decompose(type="additive") -> fit1hA 

p1<-autoplot(neg_ts10, series="Data") +   
  autolayer(trendcycle(fit10A), series="Trend") +   
  autolayer(seasadj(fit10A), series="Seasonally Adjusted") +  
  xlab("Time") + ylab("Number of tweets") +   
  ggtitle("Additive Decomposition: negative tweets every 10 minutes") +   scale_colour_manual(values=c("gray","blue","red"), breaks=c("Data","Seasonally Adjusted","Trend"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

p2<-autoplot(neg_ts30, series="Data") +   
  autolayer(trendcycle(fit30A), series="Trend") +   
  autolayer(seasadj(fit30A), series="Seasonally Adjusted") +  
  xlab("Time") + ylab("Number of tweets") +   
  ggtitle("Additive Decomposition: negative tweets every 30 minutes") +   scale_colour_manual(values=c("gray","blue","red"), breaks=c("Data","Seasonally Adjusted","Trend"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

p3<-autoplot(neg_ts1h, series="Data") +   
  autolayer(trendcycle(fit1hA), series="Trend") +   
  autolayer(seasadj(fit1hA), series="Seasonally Adjusted") +  
  xlab("Time") + ylab("Number of tweets") +   
  ggtitle("Additive Decomposition: negative tweets every hour") +   scale_colour_manual(values=c("gray","blue","red"), breaks=c("Data","Seasonally Adjusted","Trend"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

grid.arrange(p1, p2, p3, ncol=1)
```

#### Classical Multiplicative Decomposition

STL (Seasonal and Trend decomposition using Loess) is a robust, flexible decomposition method that handles any seasonality type and is resistant to outliers.

```{r}
neg_ts10 %>% decompose(type="multiplicative") -> fit10M 
neg_ts30 %>% decompose(type="multiplicative") -> fit30M 
neg_ts1h %>% decompose(type="multiplicative") -> fit1hM 

p1<-autoplot(neg_ts10, series="Data") +   
  autolayer(trendcycle(fit10M), series="Trend") +   
  autolayer(seasadj(fit10M), series="Seasonally Adjusted") +  
  xlab("Time") + ylab("Number of tweets") +   
  ggtitle("Multiplicative Decomposition: negative tweets every 10 minutes") +   scale_colour_manual(values=c("gray","blue","red"), breaks=c("Data","Seasonally Adjusted","Trend"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

p2<-autoplot(neg_ts30, series="Data") +   
  autolayer(trendcycle(fit30M), series="Trend") +   
  autolayer(seasadj(fit30M), series="Seasonally Adjusted") +  
  xlab("Time") + ylab("Number of tweets") +   
  ggtitle("Multiplicative Decomposition: negative tweets every 30 minutes") +   scale_colour_manual(values=c("gray","blue","red"), breaks=c("Data","Seasonally Adjusted","Trend"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

p3<-autoplot(neg_ts1h, series="Data") +   
  autolayer(trendcycle(fit1hM), series="Trend") +   
  autolayer(seasadj(fit1hM), series="Seasonally Adjusted") +  
  xlab("Time") + ylab("Number of tweets") +   
  ggtitle("Multiplicative Decomposition: negative tweets every hour") +   scale_colour_manual(values=c("gray","blue","red"), breaks=c("Data","Seasonally Adjusted","Trend"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

grid.arrange(p1, p2, p3, ncol=1)
```

#### Custom Decomposition: Fourier Terms + Moving Average

Since the data appears to exhibit daily periodicity, the number of Fourier terms K should be selected between 3 and 5.

```{r}
# Define a range of K values to test
K_values <- 1:10 # Test K from 1 to 10

# Function to test different values of K
K_fourier_choise <- function(serie, K_values){
  # Initialise a list to store results
  aic_values <- numeric(length(K_values))

  for (i in seq_along(K_values)) {
    K <- K_values[i]
  
    # Generate Fourier terms
    fourier_terms <- fourier(serie, K = K)
  
    # Linear model with Fourier terms
    fit <- tslm(serie ~ fourier_terms)
  
    # Save the model AIC
    aic_values[i] <- AIC(fit)
  }
  return(aic_values)
}

aic_values10 <- K_fourier_choise(neg_ts10, K_values = K_values)
aic_values30 <- K_fourier_choise(neg_ts30, K_values = K_values)
aic_values1h <- K_fourier_choise(neg_ts1h, K_values = K_values)


# Find the optimal K
best_K10 <- K_values[which.min(aic_values10)]
best_K30 <- K_values[which.min(aic_values30)]
best_K1h <- K_values[which.min(aic_values1h)]

# Print the result
cat("Best K for 10-minute frequency:", best_K10, "\n")
cat("Best K for 30-minute frequency:", best_K30, "\n")
cat("Best K for 1-hour frequency:", best_K1h, "\n")
```

```{r, include=FALSE}
# Identify periodicity 
fourier_terms10 <- fourier(neg_ts10, K = 5)
fourier_terms30 <- fourier(neg_ts30, K = 2)
fourier_terms1h <- fourier(neg_ts1h, K = 2)

F10 <- tslm(neg_ts10 ~ fourier_terms10)
F30 <- tslm(neg_ts30 ~ fourier_terms30)
F1h <- tslm(neg_ts1h ~ fourier_terms1h)

# Extract original data and fitted values 
neg_ts10_df <- data.frame(
  time = index(neg_ts10), 
  value = coredata(neg_ts10))

neg_ts30_df <- data.frame(
  time = index(neg_ts30), 
  value = coredata(neg_ts30))

neg_ts1h_df <- data.frame(
  time = index(neg_ts1h), 
  value = coredata(neg_ts1h))

neg_ts10_df$fitted <- fitted(F10)
neg_ts30_df$fitted <- fitted(F30)
neg_ts1h_df$fitted <- fitted(F1h)
```

```{r, fig.height=10}
p1<-autoplot(neg_ts10, series="Data") +   
  autolayer(ma(neg_ts10,144), series="144-MA") +   
  autolayer(fitted(F10), series = "Fourier") +   
  xlab("10m") + ylab("freq") +   
  ggtitle("Decomposition: Fourier terms (5) + Moving Average (144)\n10-minute granularity")+   
  scale_colour_manual(values=c("Data"="grey60","144-MA"="red", "Fourier"="blue"),                       breaks=c("Data","144-MA","Fourier"))+
  theme_light()

p2<-autoplot(neg_ts30, series="Data") +   
  autolayer(ma(neg_ts30,48), series="48-MA") +   
  autolayer(fitted(F30), series = "Fourier") +   
  xlab("30m") + ylab("freq") +   
  ggtitle("Decomposition: Fourier terms (2) + Moving Average (48)\n30-minute granularity")+   
  scale_colour_manual(values=c("Data"="grey60","48-MA"="red", "Fourier"="blue"),                       breaks=c("Data","48-MA","Fourier"))+
  theme_light()

p3<-autoplot(neg_ts1h, series="Data") +   
  autolayer(ma(neg_ts1h,24), series="24-MA") +   
  autolayer(fitted(F1h), series = "Fourier") +   
  xlab("1h") + ylab("freq") +   
  ggtitle("Decomposition: Fourier terms (2) + Moving Average (24)\n1-hour granularity")+   
  scale_colour_manual(values=c("Data"="grey60","24-MA"="red", "Fourier"="blue"),                       breaks=c("Data","24-MA","Fourier"))+
  theme_light()

grid.arrange(p1, p2, p3, ncol=1)
```

We now observe the complete fitted models:

-   **Model for 10-minute series**

```{r}
model_neg10 =  tslm(neg_ts10 ~ fourier_terms10 + ma(neg_ts10, 144))
summary(model_neg10)
```

-   **Model for 30-minute series**

```{r}
model_neg30 =  tslm(neg_ts30 ~ fourier_terms30 + ma(neg_ts30, 48))
summary(model_neg30)
```

-   **Model for 1-hour series**

```{r}
model_neg1h =  tslm(neg_ts1h ~ fourier_terms1h + ma(neg_ts1h, 24))
summary(model_neg1h)
```

```{r, fig.height=10}
p1<-autoplot(neg_ts10, series="Data") +  
  autolayer(fitted(model_neg10), series = "Model") +   
  xlab("10m") + ylab("freq") +   
  ggtitle("Complete Decomposition Model:\n10-minute series")+
  scale_colour_manual(values=c("Data"="grey60","Model"="red"), 
                      breaks=c("Data","Model"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

p2<-autoplot(neg_ts30, series="Data") +  
  autolayer(fitted(model_neg30), series = "Model") +   
  xlab("30m") + ylab("freq") +   
  ggtitle("Complete Decomposition Model:\n30-minute series")+
  scale_colour_manual(values=c("Data"="grey60","Model"="red"), 
                      breaks=c("Data","Model"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

p3<-autoplot(neg_ts1h, series="Data") +  
  autolayer(fitted(model_neg1h), series = "Model") +   
  xlab("1h") + ylab("freq") +   
  ggtitle("Complete Decomposition Model:\n1-hour series")+
  scale_colour_manual(values=c("Data"="grey60","Model"="red"), 
                      breaks=c("Data","Model"))+
  scale_x_continuous(
    labels = NULL # Remove x-axis labels
  )+
  theme_light()

grid.arrange(p1, p2, p3, ncol=1)
```

The additive decomposition model assumes that the components combine additively: $y_t = T_t + S_t + R_t$, where trend, seasonality, and remainder sum to the observed value. il significato di aver fatto la decomposizione.

### Model Evaluation

Evaluate the decomposition models to select the best-fitting one.

```{r, include=FALSE}
calculate_metrics <- function(series, decomp_add, decomp_mult, stl_model, model_fourier) {
  # Check that the series is not empty
  if (length(series) == 0 || all(is.na(series))) {
    stop("The time series is empty or contains only NA values.")
  }
  
  # Function to compute error metrics
  error_metrics <- function(actual, resid) {
    # Exclude NA or invalid values
    valid_indices <- !is.na(actual) & !is.na(resid) & actual != 0
    
    if (sum(valid_indices) == 0) {
      return(c(MAE = NA, RMSE = NA, MAPE = NA)) # No valid data
    }
    
    actual <- actual[valid_indices]
    resid <- resid[valid_indices]
    
    # Compute metrics
    mae <- mean(abs(resid))
    rmse <- sqrt(mean(resid^2))
    mape <- mean(abs(resid / actual)) * 100
    return(c(MAE = mae, RMSE = rmse, MAPE = mape))
  }
  
  # 1. Additive Decomposition
  if (any(series <= 0)) {
    
    metrics_add <- c(MAE = NA, RMSE = NA, MAPE = NA)
  } else if (is.null(decomp_add) || length(decomp_add$random) == 0) {
    
    metrics_add <- c(MAE = NA, RMSE = NA, MAPE = NA)
  } else {
    if (anyNA(decomp_add$random)) {
      
      resid_no_na <- na.omit(decomp_add$random)
      metrics_add <- error_metrics(series[!is.na(decomp_add$random)], resid_no_na)
    } else {
      metrics_add <- error_metrics(series, decomp_add$random)
    }
  }
  
    # 2. Multiplicative Decomposition
  if (any(series <= 0)) {
    
    metrics_mult <- c(MAE = NA, RMSE = NA, MAPE = NA)
  } else if (is.null(decomp_mult) || length(decomp_mult$random) == 0) {
    
    metrics_mult <- c(MAE = NA, RMSE = NA, MAPE = NA)
  } else {
    if (anyNA(decomp_mult$random)) {
      
      resid_no_na <- na.omit(decomp_mult$random)
      metrics_mult <- error_metrics(series[!is.na(decomp_mult$random)], resid_no_na)
    } else {
      metrics_mult <- error_metrics(series, decomp_mult$random)
    }
  }
  
  
  # 3. STL Decomposition
  if (is.null(stl_model) || ncol(stl_model$time.series) == 0) {
    
    metrics_stl <- c(MAE = NA, RMSE = NA, MAPE = NA)
  } else if (anyNA(stl_model$time.series[, "remainder"])) {
    
    resid_no_na <- na.omit(stl_model$time.series[, "remainder"])
    metrics_stl <- error_metrics(series[!is.na(stl_model$time.series[, "remainder"])], resid_no_na)
  } else {
    metrics_stl <- error_metrics(series, stl_model$time.series[, "remainder"])
  }
  
  # 4. Fourier Model
  if (is.null(model_fourier)) {
    
    metrics_fourier <- c(MAE = NA, RMSE = NA, MAPE = NA)
  } else {
    residuals_fourier <- residuals(model_fourier)
    if (anyNA(residuals_fourier)) {
      
      resid_no_na <- na.omit(residuals_fourier)
      metrics_fourier <- error_metrics(series[!is.na(residuals_fourier)], resid_no_na)
    } else {
      metrics_fourier <- error_metrics(series, residuals_fourier)
    }
  }
  
  # Final comparison metrics table
  metrics_table <- rbind(
    Additive = metrics_add,
    Multiplicative = metrics_mult,
    STL = metrics_stl,
    Fourier = metrics_fourier
  )
  
  # Return the metrics table
  return(metrics_table)
}

```

-   **Decomposition model evaluation for 10-minute granularity series**

```{r}
metrics_table10 <- calculate_metrics(
  series = neg_ts10,
  decomp_add = fit10A,
  decomp_mult = fit10M, 
  stl_model = fit10, 
  model_fourier = model_neg10
)

metrics_table10
```

-   **Decomposition model evaluation for 30-minute granularity series**

```{r}
metrics_table30 <- calculate_metrics(
  series = neg_ts30,
  decomp_add = fit30A,
  decomp_mult = fit30M, 
  stl_model = fit30, 
  model_fourier = model_neg30
)

metrics_table30
```

-   **Decomposition model evaluation for 1-hour granularity series**

```{r}
metrics_table1h <- calculate_metrics(
  series = neg_ts1h, 
  decomp_add = fit1hA,
  decomp_mult = fit1hM, 
  stl_model = fit1h, 
  model_fourier = model_neg1h
)

metrics_table1h
```
