---
title: "Time Series Analysis - Sentiment Score"
author: "Rosario Pio Gnazzo, Antonio Pauzano"
date: "2024-12-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message=FALSE, 
                      warning=FALSE, 
                      fig.align='center', fig.width = 10)
options(xts_check_TZ = FALSE)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(highfrequency)
library(xts)
library(forecast)
library(lubridate)
library(gridExtra)
library(urca)
library(tibble)
```

```{r}
# Load the dataset.
# The CSV file must be placed in the root of the repository (one level above this script).
# See data/README.md for instructions on how to obtain the dataset.
Sentiment_fr_tweet_2023 <- read_csv2("../Sentiment_fr_tweet_2023.csv")

dataset <- Sentiment_fr_tweet_2023

dataset <- dataset %>%
  mutate(userid = as.character(userid),
         following = as.numeric(following),
         followers = as.numeric(followers),
         totaltweets = as.numeric(totaltweets),
         tweetid = as.character(tweetid),
         retweetcount = as.numeric(retweetcount),
         favorite_count = as.numeric(favorite_count),
         is_retweet = as.factor(ifelse(is_retweet == FALSE, 0, 1)),
         is_quote_status = as.factor(ifelse(is_quote_status == FALSE, 0, 1)),
         sentiment = as.factor(sentiment),
         score = as.numeric(score),
         score_std = scale(score))
```

# Time Series Analysis of the Numerical Feature *Score*

We observe how the *score* variable changes over time for the three *sentiment* classes (neg, neu, and pos):

```{r, fig.width=15}
df <- dataset %>% select(tweetcreatedts, score, sentiment)

ggplot(df, aes(x = tweetcreatedts, y = score, color = sentiment)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Time series of the <Score> variable") +
  facet_wrap(~sentiment, scales = "free", ncol = 1) +
  theme_light()
```

We observe their distribution:

```{r}
ggplot(df, aes(x = score, fill = sentiment)) +
  geom_density(alpha = 0.4) +
  scale_fill_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Score Density by Sentiment") +
  theme_light() +
  labs(fill = "Sentiment", x = "Score", y = "Density")
```

We observe how the series changes when a *Box-Cox* transformation is applied to mitigate data variability:

```{r, fig.height=5, include=FALSE}
lambdaNeg <- df %>% filter(sentiment=="neg") %>% select(score) %>% BoxCox.lambda()
lambdaNeu <- df %>% filter(sentiment=="neu") %>% select(score) %>% BoxCox.lambda()
lambdaPos <- df %>% filter(sentiment=="pos") %>% select(score) %>% BoxCox.lambda()
#sono tutti uguali

ggplot(df, aes(x = tweetcreatedts, y = BoxCox(score, lambdaNeg), color = sentiment)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Time series of the <Score> variable - post Box-Cox transformation")+
  facet_wrap(~sentiment, scales = "free", ncol = 1)+
  theme_light()
```

The transformation changes very little, so we will continue the analysis using the original series.

## Stationarity and Autocorrelation Analysis of the Time Series

The series most likely show no strong Trend, Periodicity, or Cyclicity. Given their behaviour, they likely follow a **White Noise** distribution and are **stationary**, remaining stable around the mean with constant variance over time.

To verify this, we use two hypothesis tests:

1.  **Ljung-Box Test**

    -   *Purpose*: Test for **residual autocorrelation** in a time series

    -   *Null hypothesis (H0)*: No autocorrelation up to a given lag -> *white noise*

    -   *Alternative hypothesis (H1)*: Autocorrelation exists up to a given lag

    -   Conclusion:

        -   A small *p-value* (\< 0.05) suggests rejecting H0, indicating significant autocorrelation.

        -   A large *p-value* indicates no evidence of autocorrelation.

2.  **Kwiatkowski-Phillips-Schmidt-Shin (KPSS) Test**

    -   *Purpose*: Test whether a time series is **stationary**.

    -   *Null hypothesis (H0)*: The series is stationary.

    -   *Alternative hypothesis (H1)*: The series is not stationary.

    -   Conclusion:

        -   A small *p-value* (e.g., \< 0.05) suggests rejecting H0, meaning the series is non-stationary and may require differencing.

        -   A large *p-value* fails to reject H0, suggesting stationary behaviour.

For the negative score time series:

```{r, include=FALSE}
# Extract time series
negTS <- df %>% filter(sentiment=="neg") %>% select(score, tweetcreatedts)
neuTS <- df %>% filter(sentiment=="neu") %>% select(score, tweetcreatedts)
posTS <- df %>% filter(sentiment=="pos") %>% select(score, tweetcreatedts)
# Test for the negative score series
Box.test(negTS$score, lag = 10, type = "Ljung-Box")
negTS %>% select(score) %>% as.ts() %>% ur.kpss() %>% summary()
# Test for the neutral score series
Box.test(neuTS$score, lag = 10, type = "Ljung-Box")
neuTS %>% select(score) %>% as.ts() %>% ur.kpss() %>% summary()
# Test for the positive score series
Box.test(posTS$score, lag = 10, type = "Ljung-Box")
posTS %>% select(score) %>% as.ts() %>% ur.kpss() %>% summary()
```

+---------------+----------------------------+---------------------------+
| Score         | KPSS Test                  | Ljung-Box Test            |
+:=============:+:==========================:+:=========================:+
| Negative      | ```                        | ```                       |
|               | test-statistic is: 73.8069 | p-value < 2.2e-16         |
|               | ```                        | ```                       |
|               |                            |                           |
|               | Non-stationary            | Autocorrelation present |
+---------------+----------------------------+---------------------------+
| Neutral       | ```                        | ```                       |
|               | test-statistic is: 17.6576 | p-value < 2.2e-16         |
|               | ```                        | ```                       |
|               |                            |                           |
|               | Non-stationary            | Autocorrelation present |
+---------------+----------------------------+---------------------------+
| Positive      | ```                        | ```                       |
|               | test-statistic is: 34.624  | p-value < 2.2e-16         |
|               | ```                        | ```                       |
|               |                            |                           |
|               | Non-stationary            | Autocorrelation present |
+---------------+----------------------------+---------------------------+

The test results show that the series **are not stationary** and do not follow a white noise distribution, meaning they **exhibit autocorrelation** at the defined lags. Note that *autocorrelation does not imply non-stationarity* and vice versa: stationarity concerns general time-level properties of the series, while the Ljung-Box test focuses on temporal dependence patterns over specific lag sub-intervals.

We examine whether this structure persists as we reduce the observation frequency. We focus on frequencies of 10 minutes, 30 minutes, and 1 hour.

### Analysis at **10-minute** frequency:

```{r, include=FALSE}
##### visualise ts aggregated to 10-minute intervals #####
negTS <- xts(negTS$score, order.by = negTS$tweetcreatedts)
neuTS <- xts(neuTS$score, order.by = neuTS$tweetcreatedts)
posTS <- xts(posTS$score, order.by = posTS$tweetcreatedts)
# Aggregate data to 10-minute observation frequency
score10neg <- aggregateTS(negTS, alignBy = "minutes", alignPeriod = 10)
score10neu <- aggregateTS(neuTS, alignBy = "minutes", alignPeriod = 10)
score10pos <- aggregateTS(posTS, alignBy = "minutes", alignPeriod = 10)

dfTS10 <- data.frame(
  time = index(score10neg),
  neg = as.numeric(score10neg),
  neu = as.numeric(score10neu),
  pos = as.numeric(score10pos)
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

# Create dataset for predominant sentiment
predominant_sentiment <- data.frame(
  time = index(score10neg),
  neg = as.numeric(score10neg),
  neu = as.numeric(score10neu),
  pos = as.numeric(score10pos)
) %>%
  rowwise() %>%
  mutate(
    predominant = c("neg", "neu", "pos")[which.max(c(neg, neu, pos))],
    score = max(c(neg, neu, pos), na.rm = TRUE)
  ) %>%
  select(time, predominant, score)

# Compute sentiment frequencies over time
sentiment_frequencies <- predominant_sentiment %>%
  group_by(predominant) %>%
  summarise(frequency = n())

# Compute first-derivative series
deriv_neg10 <- diff(score10neg) / diff(index(score10neg))
deriv_neu10 <- diff(score10neu) / diff(index(score10neu))
deriv_pos10 <- diff(score10pos) / diff(index(score10pos))

# Create data frame for plotting
df_deriv10 <- data.frame(
  time = index(deriv_neg10),
  neg = as.numeric(deriv_neg10),
  neu = as.numeric(deriv_neu10),
  pos = as.numeric(deriv_pos10)
) %>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "Derivative")

# Compute polarization index
polarization_index <- data.frame(
  time = index(score10neg),
  neg = as.numeric(score10neg),
  neu = as.numeric(score10neu),
  pos = as.numeric(score10pos)
) %>%
  rowwise() %>%
  mutate(
    max_score = max(neg, neu, pos, na.rm = TRUE),
    mean_other = mean(c(neg, neu, pos)[-which.max(c(neg, neu, pos))], na.rm = TRUE),
    polarization = max_score - mean_other
  ) %>%
  select(time, polarization)
```

```{r}
#PLOT
plot <- ggplot(dfTS10, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Score Series")+
  facet_wrap(~Series, scales = "free", ncol = 1)+
  theme_light() +
  theme(legend.position = "top")

# plot the derivatives
plot_derivate <- ggplot(df_deriv10, aes(x = time, y = Derivative, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Temporal derivatives of the Score series") +
  facet_wrap(~Series, scales = "free", ncol = 1) +
  theme_light() +
  theme(legend.position = "top") +
  labs(color = "Sentiment", x = "Time", y = "Score Derivative")

grid.arrange(plot, plot_derivate, ncol = 2)






# # Creazione della matrice di transizione
# transition_matrix <- table(predominant_sentiment$predominant[-nrow(predominant_sentiment)],
#                            predominant_sentiment$predominant[-1])
# 
# # Compute transition probabilities
# transition_probabilities <- prop.table(transition_matrix, 1)
# 
# # Visualizzazione della matrice di transizione come heatmap
# ggplot(as.data.frame(as.table(transition_probabilities)), aes(Var1, Var2, fill = Freq)) +
#   geom_tile() +
#   scale_fill_gradient2(low = "white", high = "blue", mid = "red", midpoint = 0.5) +
#   labs(title = "Matrice di Transizione tra Sentiment", x = "Sentiment Successivo", y = "Sentiment Precedente") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

We observe the stationarity and autocorrelation test results for the series:

```{r, include=FALSE}
Box.test(score10neg, lag = 10, type = "Ljung-Box")
score10neg %>% ur.kpss() %>% summary()

Box.test(score10neu, lag = 10, type = "Ljung-Box")
score10neu %>% ur.kpss() %>% summary()

Box.test(score10pos, lag = 10, type = "Ljung-Box")
score10pos %>% ur.kpss() %>% summary()



### Verify stationarity and autocorrelation of derivative series:
Box.test(deriv_neg10, lag = 10, type = "Ljung-Box")
deriv_neg10 %>% as.ts() %>% ur.kpss() %>% summary()

Box.test(deriv_neu10, lag = 10, type = "Ljung-Box")
deriv_neu10 %>% as.ts() %>% ur.kpss() %>% summary()

Box.test(deriv_pos10, lag = 10, type = "Ljung-Box")
deriv_pos10 %>% as.ts() %>% ur.kpss() %>% summary()
```

+--------------------------+--------------------------------------+------------------------------+
|                          | KPSS Test                            | Ljung-Box Test               |
+:========================:+:====================================:+:============================:+
| **Score Series**         | test-statistic is: **1.6076** (N.S.) | p-value = **0.5361** (N.A.)  |
|                          |                                      |                              |
|                          | test-statistic is: **1.0155** (N.S.) | p-value = **3.389e-06** (A.) |
|                          |                                      |                              |
|                          | test-statistic is: **1.5981** (N.S.) | p-value = **0.01512** (A.)   |
+--------------------------+--------------------------------------+------------------------------+
| **Score Derivative**     | test-statistic is: **0.0111** (S.)   | p-value \< **2.2e-16** (A.)  |
|                          |                                      |                              |
|                          | test-statistic is: **0.006** (S.)    | p-value \< **2.2e-16** (A.)  |
|                          |                                      |                              |
|                          | test-statistic is: **0.0068** (S.)   | p-value \< **2.2e-16** (A.)  |
+--------------------------+--------------------------------------+------------------------------+

At this frequency, the test statistic for the stationarity hypothesis (untransformed series) has decreased substantially but remains above the 1% critical threshold. The temporal derivative series are already stationary at this level. Only the negative score series becomes White Noise and loses autocorrelation at this frequency.

We observe the overall score behaviour at this frequency:

```{r, fig.height=5, fig.width=10}
# Plot del sentiment predominante
plot_predominant <- ggplot(predominant_sentiment, aes(x = time, y = score, color = predominant)) +
  geom_point() +  # Usa geom_point per visualizzare ogni massimo
  geom_line(aes(group = 1)) +  # Connette i punti in ordine temporale
  scale_color_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Predominant sentiment every 10 minutes") +
  theme_light() +
  theme(legend.position = "top") +
  labs(color = "Sentiment", x = "Tempo", y = "Score")

# Plot delle frequenze dei sentiment
plot_frequencies <- ggplot(sentiment_frequencies, aes(x = predominant, y = frequency, fill = predominant)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  theme_light() +
  theme(legend.position = "top") +
  labs(fill = "Sentiment", x = "Sentiment", y = "Frequency")

# # Plot dell'indice di polarizzazione
# plot_polarizzazione <-ggplot(polarization_index, aes(x = time, y = polarization)) +
#   geom_line(color = "purple") +
#   ggtitle("Score Polarization Index") +
#   theme_light() +
#   labs(x = "Time", y = "Polarization Index")

grid.arrange(plot_predominant, 
             plot_frequencies,
             # plot_polarizzazione, 
             ncol = 1)
```

### Analysis at **30-minute** frequency:

```{r, include=FALSE}
##### visualise ts aggregated to 30-minute intervals #####
score30neg <- aggregateTS(negTS, alignBy = "minutes", alignPeriod = 30)
score30neu <- aggregateTS(neuTS, alignBy = "minutes", alignPeriod = 30)
score30pos <- aggregateTS(posTS, alignBy = "minutes", alignPeriod = 30)

dfTS30 <- data.frame(
  time = index(score30neg),
  neg = as.numeric(score30neg),
  neu = as.numeric(score30neu),
  pos = as.numeric(score30pos)
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

# Create dataset for predominant sentiment
predominant_sentiment <- data.frame(
  time = index(score30neg),
  neg = as.numeric(score30neg),
  neu = as.numeric(score30neu),
  pos = as.numeric(score30pos)
) %>%
  rowwise() %>%
  mutate(
    predominant = c("neg", "neu", "pos")[which.max(c(neg, neu, pos))],
    score = max(c(neg, neu, pos), na.rm = TRUE)
  ) %>%
  select(time, predominant, score)

# Compute sentiment frequencies over time
sentiment_frequencies <- predominant_sentiment %>%
  group_by(predominant) %>%
  summarise(frequency = n())

# Compute derivatives
deriv_neg30 <- diff(score30neg) / diff(index(score30neg))
deriv_neu30 <- diff(score30neu) / diff(index(score30neu))
deriv_pos30 <- diff(score30pos) / diff(index(score30pos))

# Create data frame for plotting
df_deriv30 <- data.frame(
  time = index(deriv_neg30),
  neg = as.numeric(deriv_neg30),
  neu = as.numeric(deriv_neu30),
  pos = as.numeric(deriv_pos30)
) %>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "Derivative")

# Compute polarization index
polarization_index <- data.frame(
  time = index(score30neg),
  neg = as.numeric(score30neg),
  neu = as.numeric(score30neu),
  pos = as.numeric(score30pos)
) %>%
  rowwise() %>%
  mutate(
    max_score = max(neg, neu, pos, na.rm = TRUE),
    mean_other = mean(c(neg, neu, pos)[-which.max(c(neg, neu, pos))], na.rm = TRUE),
    polarization = max_score - mean_other
  ) %>%
  select(time, polarization)
```

```{r}
plot<-ggplot(dfTS30, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Score Series")+
  facet_wrap(~Series, scales = "free", ncol = 1)+
  theme_light() +
  theme(legend.position = "top")

# plot the derivatives
plot_derivate <- ggplot(df_deriv30, aes(x = time, y = Derivative, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Temporal derivatives of the Score series") +
  facet_wrap(~Series, scales = "free", ncol = 1)+
  theme_light() +
  theme(legend.position = "top") +
  labs(color = "Sentiment", x = "Time", y = "Score Derivative")

grid.arrange(plot, plot_derivate, ncol=2)
```

We observe the test results:

```{r, include=FALSE}
Box.test(score30neg, lag = 10, type = "Ljung-Box")
score30neg %>% ur.kpss() %>% summary()

Box.test(score30neu, lag = 10, type = "Ljung-Box")
score30neu %>% ur.kpss() %>% summary()

Box.test(score30pos, lag = 10, type = "Ljung-Box")
score30pos %>% ur.kpss() %>% summary()

### Verify stationarity and autocorrelation of derivative series:
Box.test(deriv_neg30, lag = 10, type = "Ljung-Box")
deriv_neg30 %>% as.ts() %>% ur.kpss() %>% summary()

Box.test(deriv_neu30, lag = 10, type = "Ljung-Box")
deriv_neu30 %>% as.ts() %>% ur.kpss() %>% summary()

Box.test(deriv_pos30, lag = 10, type = "Ljung-Box")
deriv_pos30 %>% as.ts() %>% ur.kpss() %>% summary()
```

+--------------------------+--------------------------------------+------------------------------+
| Score                    | KPSS Test                            | Ljung-Box Test               |
+:========================:+:====================================:+:============================:+
| **Score Series**         | test-statistic is: **0.7737** (N.S.) | p-value = **0.07077** (N.A.) |
|                          |                                      |                              |
|                          | test-statistic is: **0.2137** (S.)   | p-value = **0.0007944** (A.) |
|                          |                                      |                              |
|                          | test-statistic is: **0.2168** (S.)   | p-value = **0.357** (N.A.)   |
+--------------------------+--------------------------------------+------------------------------+
| **Score Derivatives**    | test-statistic is: **0.0111** (S.)   | p-value \< **2.2e-16** (A.)  |
|                          |                                      |                              |
|                          | test-statistic is: **0.006** (S.)    | p-value = **3.563e-13** (A.) |
|                          |                                      |                              |
|                          | test-statistic is: **0.0068** (S.)   | p-value = **1.984e-12** (A.) |
+--------------------------+--------------------------------------+------------------------------+

```{r, fig.height=5, fig.width=6}
# Plot del sentiment predominante
plot_predominant <- ggplot(predominant_sentiment, aes(x = time, y = score, color = predominant)) +
  geom_point() +  # Usa geom_point per visualizzare ogni massimo
  geom_line(aes(group = 1)) +  # Connette i punti in ordine temporale
  scale_color_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Predominant sentiment every 30 minutes") +
  theme_light() +
  theme(legend.position = "top") +
  labs(color = "Sentiment", x = "Tempo", y = "Score")

# Plot delle frequenze dei sentiment
plot_frequencies <- ggplot(sentiment_frequencies, aes(x = predominant, y = frequency, fill = predominant)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  theme_light() +
  theme(legend.position = "top") +
  labs(fill = "Sentiment", x = "Sentiment", y = "Frequency")

# # Plot dell'indice di polarizzazione
# plot_polarizzazione <-ggplot(polarization_index, aes(x = time, y = polarization)) +
#   geom_line(color = "purple") +
#   ggtitle("Score Polarization Index") +
#   theme_light() +
#   labs(x = "Time", y = "Polarization Index")

grid.arrange(plot_predominant, 
             plot_frequencies,
             # plot_polarizzazione, 
             ncol = 1)
```

### Analysis at **1-hour** frequency:

```{r}
##### visualise ts aggregated to 1-hour intervals #####
score1hneg <- aggregateTS(negTS, alignBy = "hours", alignPeriod = 1)
score1hneu <- aggregateTS(neuTS, alignBy = "hours", alignPeriod = 1)
score1hpos <- aggregateTS(posTS, alignBy = "hours", alignPeriod = 1)

dfTS1h <- data.frame(
  time = index(score1hneg),
  neg = as.numeric(score1hneg),
  neu = as.numeric(score1hneu),
  pos = as.numeric(score1hpos)
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

# Create dataset for predominant sentiment
predominant_sentiment <- data.frame(
  time = index(score1hneg),
  neg = as.numeric(score1hneg),
  neu = as.numeric(score1hneu),
  pos = as.numeric(score1hpos)
) %>%
  rowwise() %>%
  mutate(
    predominant = c("neg", "neu", "pos")[which.max(c(neg, neu, pos))],
    score = max(c(neg, neu, pos), na.rm = TRUE)
  ) %>%
  select(time, predominant, score)

# Compute sentiment frequencies over time
sentiment_frequencies <- predominant_sentiment %>%
  group_by(predominant) %>%
  summarise(frequency = n())

# Compute derivatives
deriv_neg1h <- diff(score1hneg) / diff(index(score1hneg))
deriv_neu1h <- diff(score1hneu) / diff(index(score1hneu))
deriv_pos1h <- diff(score1hpos) / diff(index(score1hpos))

# Create data frame for plotting
df_deriv1h <- data.frame(
  time = index(deriv_neg1h),
  neg = as.numeric(deriv_neg1h),
  neu = as.numeric(deriv_neu1h),
  pos = as.numeric(deriv_pos1h)
) %>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "Derivative")

# Compute polarization index
polarization_index <- data.frame(
  time = index(score1hneg),
  neg = as.numeric(score1hneg),
  neu = as.numeric(score1hneu),
  pos = as.numeric(score1hpos)
) %>%
  rowwise() %>%
  mutate(
    max_score = max(neg, neu, pos, na.rm = TRUE),
    mean_other = mean(c(neg, neu, pos)[-which.max(c(neg, neu, pos))], na.rm = TRUE),
    polarization = max_score - mean_other
  ) %>%
  select(time, polarization)
```

```{r}
plot<-ggplot(dfTS1h, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Score Series")+
  facet_wrap(~Series, scales = "free", ncol = 1)+
  theme_light() +
  theme(legend.position = "top")
# plot the derivatives
plot_derivate <- ggplot(df_deriv1h, aes(x = time, y = Derivative, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Temporal derivatives of the Score series") +
  facet_wrap(~Series, scales = "free", ncol = 1)+
  theme_light() +
  theme(legend.position = "top") +
  labs(color = "Sentiment", x = "Time", y = "Score Derivative")

grid.arrange(plot, plot_derivate, ncol=2)
```

We observe the test results:

```{r, include=FALSE}
Box.test(score1hneg, lag = 10, type = "Ljung-Box")
score1hneg %>% ur.kpss() %>% summary()

Box.test(score1hneu, lag = 10, type = "Ljung-Box")
score1hneu %>% ur.kpss() %>% summary()

Box.test(score1hpos, lag = 10, type = "Ljung-Box")
score1hpos %>% ur.kpss() %>% summary()

### Verify stationarity and autocorrelation of derivative series:
Box.test(deriv_neg1h, lag = 10, type = "Ljung-Box")
deriv_neg1h %>% as.ts() %>% ur.kpss() %>% summary()

Box.test(deriv_neu1h, lag = 10, type = "Ljung-Box")
deriv_neu1h %>% as.ts() %>% ur.kpss() %>% summary()

Box.test(deriv_pos1h, lag = 10, type = "Ljung-Box")
deriv_pos1h %>% as.ts() %>% ur.kpss() %>% summary()
```

+--------------------------+------------------------------------+------------------------------+
| Score                    | KPSS Test                          | Ljung-Box Test               |
+:========================:+:==================================:+:============================:+
| **Score Series**         | test-statistic is: **0.3793** (S.) | p-value = **0.1761** (N.A.)  |
|                          |                                    |                              |
|                          | test-statistic is: **0.1504** (S.) | p-value = **0.09925** (N.A.) |
|                          |                                    |                              |
|                          | test-statistic is: **0.1444** (S.) | p-value = **0.4608** (N.A.)  |
+--------------------------+------------------------------------+------------------------------+
| **Score Derivatives**    | test-statistic is: **0.0111** (S.) | p-value = **2.32e-05** (A.)  |
|                          |                                    |                              |
|                          | test-statistic is: **0.006** (S.)  | p-value = **4.77e-06** (A.)  |
|                          |                                    |                              |
|                          | test-statistic is: **0.0068** (S.) | p-value = **1.427e-11** (A.) |
+--------------------------+------------------------------------+------------------------------+

```{r, fig.height=5, fig.width=6}
# Plot del sentiment predominante
plot_predominant <- ggplot(predominant_sentiment, aes(x = time, y = score, color = predominant)) +
  geom_point() +  # Usa geom_point per visualizzare ogni massimo
  geom_line(aes(group = 1)) +  # Connette i punti in ordine temporale
  scale_color_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  ggtitle("Predominant sentiment every hour") +
  theme_light() +
  theme(legend.position = "top") +
  labs(color = "Sentiment", x = "Tempo", y = "Score")

# Plot delle frequenze dei sentiment
plot_frequencies <- ggplot(sentiment_frequencies, aes(x = predominant, y = frequency, fill = predominant)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("neg" = "darkred", "neu" = "darkblue", "pos" = "darkgreen")) +
  theme_light() +
  theme(legend.position = "top") +
  labs(fill = "Sentiment", x = "Sentiment", y = "Frequency")

# # Plot dell'indice di polarizzazione
# plot_polarizzazione <-ggplot(polarization_index, aes(x = time, y = polarization)) +
#   geom_line(color = "purple") +
#   ggtitle("Score Polarization Index") +
#   theme_light() +
#   labs(x = "Time", y = "Polarization Index")

grid.arrange(plot_predominant, 
             plot_frequencies, 
             # plot_polarizzazione, 
             ncol = 1)
```

#### Observations

For the Score series, as we aggregate to lower frequencies the series tend towards a *white noise* distribution; stationarity is achieved by all series at the lowest frequency. The temporal derivative series, by contrast, consistently exhibit stationarity and autocorrelation regardless of the frequency level.

## Building Statistical Models

### Non-Seasonal ARIMA Models

ARIMA stands for AutoRegressive Integrated Moving Average. It combines the concept of differencing with autoregressive and moving-average components.

$$
y'_{t} = c + \phi_{1}y'_{t-1} + \dots + \phi_{p}y'_{t-p} + \theta_{1}\varepsilon_{t-1} + \dots + \theta_{p}\varepsilon_{t-p}
$$

In *backshift notation* this becomes:

$$
(1 - \phi_{1}B - \dots - \phi_{p}B^{p})(1-B)^{d}y_{t} = c + (1+\theta_{1}B+\dots+\theta_{q}B^{q})\varepsilon_{t}
$$

Where $y'_{t}$ is the differenced series (potentially differenced more than once), and the predictors are the lagged values and lagged errors of the series. This structure is called the **ARIMA(p, d, q)** **Model**, where

-   *p* = order of the autoregressive part
-   *d* = degree of differencing
-   *q* = order of the moving average

Many well-known models are special cases of ARIMA(p, d, q):

-   White Noise = ARIMA(0, 0 , 0)

-   Random Walk = ARIMA(0, 1, 0)

-   Autoregression = ARIMA(p, 0, 0)

-   Moving Average = ARIMA(0, 0, q)

We fit an ARIMA(p, d, q) model to each of the score series and their temporal derivatives.

```{r}
arima_summary <- function(model, name) {
  tibble(
    Serie = name,
    ARIMA = paste0("(", model$arma[1], ",", model$arma[6], ",", model$arma[2], ")"),
    AIC = model$aic,
    AICc = model$aicc,
    BIC = model$bic
  )
}
```

```{r, include=FALSE}
## 10 minutes ##
fitNEG10 <- auto.arima(score10neg, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitNEU10 <- auto.arima(score10neu, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitPOS10 <- auto.arima(score10pos, seasonal = FALSE, 
                     stepwise = FALSE, 
                     approximation = FALSE)

## 30 minutes ##
fitNEG30 <- auto.arima(score30neg, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitNEU30 <- auto.arima(score30neu, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitPOS30 <- auto.arima(score30pos, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

## 1 hour ##
fitNEG1h <- auto.arima(score1hneg, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitNEU1h <- auto.arima(score1hneu, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitPOS1h <- auto.arima(score1hpos, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

results <- bind_rows(
  arima_summary(fitNEG10, "10min_NEG"),
  arima_summary(fitNEU10, "10min_NEU"),
  arima_summary(fitPOS10, "10min_POS"),
  arima_summary(fitNEG30, "30min_NEG"),
  arima_summary(fitNEU30, "30min_NEU"),
  arima_summary(fitPOS30, "30min_POS"),
  arima_summary(fitNEG1h, "1h_NEG"),
  arima_summary(fitNEU1h, "1h_NEU"),
  arima_summary(fitPOS1h, "1h_POS")
)

#### SERIE DELLE DERIVATE TEMPORALI ####
## 10 minutes ##
fitNEG10d <- auto.arima(deriv_neg10, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitNEU10d <- auto.arima(deriv_neu10, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitPOS10d <- auto.arima(deriv_pos10, seasonal = FALSE, 
                     stepwise = FALSE, 
                     approximation = FALSE)

## 30 minutes ##
fitNEG30d <- auto.arima(deriv_neg30, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitNEU30d <- auto.arima(deriv_neu30, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitPOS30d <- auto.arima(deriv_pos30, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

## 1 hour ##
fitNEG1hd <- auto.arima(deriv_neg1h, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitNEU1hd <- auto.arima(deriv_neu1h, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

fitPOS1hd <- auto.arima(deriv_pos1h, seasonal = FALSE, stepwise = FALSE, 
                     approximation = FALSE)

resultsd <- bind_rows(
  arima_summary(fitNEG10d, "10min_NEG"),
  arima_summary(fitNEU10d, "10min_NEU"),
  arima_summary(fitPOS10d, "10min_POS"),
  arima_summary(fitNEG30d, "30min_NEG"),
  arima_summary(fitNEU30d, "30min_NEU"),
  arima_summary(fitPOS30d, "30min_POS"),
  arima_summary(fitNEG1hd, "1h_NEG"),
  arima_summary(fitNEU1hd, "1h_NEU"),
  arima_summary(fitPOS1hd, "1h_POS")
)

print(results)
print(resultsd)
```

```{r}
r10 <- bind_rows(
  arima_summary(fitNEG10, "10min_NEG"),
  arima_summary(fitNEU10, "10min_NEU"),
  arima_summary(fitPOS10, "10min_POS"),
  arima_summary(fitNEG10d, "10min_NEG_Deriv"),
  arima_summary(fitNEU10d, "10min_NEU_Deriv"),
  arima_summary(fitPOS10d, "10min_POS_Deriv")
)
r30 <- bind_rows(
  arima_summary(fitNEG30, "30min_NEG"),
  arima_summary(fitNEU30, "30min_NEU"),
  arima_summary(fitPOS30, "30min_POS"),
  arima_summary(fitNEG30d, "30min_NEG_Deriv"),
  arima_summary(fitNEU30d, "30min_NEU_Deriv"),
  arima_summary(fitPOS30d, "30min_POS_Deriv")
)
r1h <- bind_rows(
  arima_summary(fitNEG1h, "1h_NEG"),
  arima_summary(fitNEU1h, "1h_NEU"),
  arima_summary(fitPOS1h, "1h_POS"),
  arima_summary(fitNEG1hd, "1h_NEG_Deriv"),
  arima_summary(fitNEU1hd, "1h_NEU_Deriv"),
  arima_summary(fitPOS1hd, "1h_POS_Deriv")
)

print(r10)
print(r30)
print(r1h)
```

#### Score Time Series

We observe how the ARIMA models fit the data:

-   **Frequenza di 10 minuti**

```{r, fig.height=8}
## 10 minutes ##
#fitNEG10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df10NEG <- data.frame(
  time = index(score10neg),
  neg = as.numeric(score10neg),
  fittedNEG = as.numeric(fitted(fitNEG10))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(df10NEG, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "fittedNEG" = "darkred")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

#fitNEU10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df10NEU <- data.frame(
  time = index(score10neu),
  neu = as.numeric(score10neu),
  fittedNEU = as.numeric(fitted(fitNEU10))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(df10NEU, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "fittedNEU" = "darkblue")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

#fitPOS10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df10POS <- data.frame(
  time = index(score10pos),
  pos = as.numeric(score10pos),
  fittedPOS = as.numeric(fitted(fitPOS10))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(df10POS, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "fittedPOS" = "darkgreen")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

grid.arrange(p1, p2, p3, ncol = 1)
```

-   **Frequenza di 30 minuti**

```{r, fig.height=8}
## 30 minutes ##
#fitNEG30 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df30NEG <- data.frame(
  time = index(score30neg),
  neg = as.numeric(score30neg),
  fittedNEG = as.numeric(fitted(fitNEG30))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(df30NEG, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "fittedNEG" = "darkred")) +
  ggtitle("ARIMA(0,1,4)")+
  theme_light()

#fitNEU30 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df30NEU <- data.frame(
  time = index(score30neu),
  neu = as.numeric(score30neu),
  fittedNEU = as.numeric(fitted(fitNEU30))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(df30NEU, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "fittedNEU" = "darkblue")) +
  ggtitle("ARIMA(1,0,1) with non-zero mean ")+
  theme_light()

#fitPOS30 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df30POS <- data.frame(
  time = index(score30pos),
  pos = as.numeric(score30pos),
  fittedPOS = as.numeric(fitted(fitPOS30))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(df30POS, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "fittedPOS" = "darkgreen")) +
  ggtitle("ARIMA(1,0,0) with non-zero mean ")+
  theme_light()

grid.arrange(p1, p2, p3, ncol = 1)
```

-   **Frequenza di 1 ora**

```{r, fig.height=8}
## 1 hour ##
#fitNEG1h %>% forecast(h=12) %>% autoplot()+theme_light()
df1hNEG <- data.frame(
  time = index(score1hneg),
  neg = as.numeric(score1hneg),
  fittedNEG = as.numeric(fitted(fitNEG1h))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(df1hNEG, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "fittedNEG" = "darkred")) +
  ggtitle("ARIMA(1,0,1) with non-zero mean")+
  theme_light()

#fitNEU1h %>% forecast(h=12) %>% autoplot()+theme_light()
df1hNEU <- data.frame(
  time = index(score1hneu),
  neu = as.numeric(score1hneu),
  fittedNEU = as.numeric(fitted(fitNEU1h))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(df1hNEU, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "fittedNEU" = "darkblue")) +
  ggtitle("ARIMA(1,0,1) with non-zero mean")+
  theme_light()

#fitPOS1h %>% forecast(h=12) %>% autoplot()+theme_light()
df1hPOS <- data.frame(
  time = index(score1hpos),
  pos = as.numeric(score1hpos),
  fittedPOS = as.numeric(fitted(fitPOS1h))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(df1hPOS, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "fittedPOS" = "darkgreen")) +
  ggtitle("ARIMA(2,0,0) with non-zero mean")+
  theme_light()

grid.arrange(p1, p2, p3, ncol = 1)
```

#### Temporal Derivative Time Series of the Score

-   **Frequenza di 10 minuti**

```{r, fig.height=8}
## 10 minutes ##
#fitNEG10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df10NEGd <- data.frame(
  time = index(deriv_neg10[2:length(deriv_neg10)]),
  neg = as.numeric(deriv_neg10[2:length(deriv_neg10)]),
  fittedNEG = as.numeric(fitted(fitNEG10d))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(df10NEGd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "fittedNEG" = "darkred")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

#fitNEU10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df10NEUd <- data.frame(
  time = index(deriv_neu10[2:length(deriv_neu10)]),
  neu = as.numeric(deriv_neu10[2:length(deriv_neu10)]),
  fittedNEU = as.numeric(fitted(fitNEU10d))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(df10NEUd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "fittedNEU" = "darkblue")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

#fitPOS10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df10POSd <- data.frame(
  time = index(deriv_pos10[2:length(deriv_pos10)]),
  pos = as.numeric(deriv_pos10[2:length(deriv_pos10)]),
  fittedPOS = as.numeric(fitted(fitPOS10d))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(df10POSd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "fittedPOS" = "darkgreen")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

grid.arrange(p1, p2, p3, ncol = 1)
```

-   **Frequenza di 30 minuti**

```{r, fig.height=8}
## 30 minutes ##
#fitNEG10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df30NEGd <- data.frame(
  time = index(deriv_neg30[2:length(deriv_neg30)]),
  neg = as.numeric(deriv_neg30[2:length(deriv_neg30)]),
  fittedNEG = as.numeric(fitted(fitNEG30d))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(df30NEGd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "fittedNEG" = "darkred")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

#fitNEU10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df30NEUd <- data.frame(
  time = index(deriv_neu30[2:length(deriv_neu30)]),
  neu = as.numeric(deriv_neu30[2:length(deriv_neu30)]),
  fittedNEU = as.numeric(fitted(fitNEU30d))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(df30NEUd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "fittedNEU" = "darkblue")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

#fitPOS10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df30POSd <- data.frame(
  time = index(deriv_pos30[2:length(deriv_pos30)]),
  pos = as.numeric(deriv_pos30[2:length(deriv_pos30)]),
  fittedPOS = as.numeric(fitted(fitPOS30d))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(df30POSd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "fittedPOS" = "darkgreen")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

grid.arrange(p1, p2, p3, ncol = 1)
```

-   **Frequenza di 1 ora**

```{r, fig.height=8}
## 1 hour ##
#fitNEG10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df1hNEGd <- data.frame(
  time = index(deriv_neg1h[2:length(deriv_neg1h)]),
  neg = as.numeric(deriv_neg1h[2:length(deriv_neg1h)]),
  fittedNEG = as.numeric(fitted(fitNEG1hd))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(df1hNEGd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "fittedNEG" = "darkred")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

#fitNEU10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df1hNEUd <- data.frame(
  time = index(deriv_neu1h[2:length(deriv_neu1h)]),
  neu = as.numeric(deriv_neu1h[2:length(deriv_neu1h)]),
  fittedNEU = as.numeric(fitted(fitNEU1hd))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(df1hNEUd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "fittedNEU" = "darkblue")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

#fitPOS10 %>% forecast(h=12) %>% autoplot(include=100)+theme_light()
df1hPOSd <- data.frame(
  time = index(deriv_pos1h[2:length(deriv_pos1h)]),
  pos = as.numeric(deriv_pos1h[2:length(deriv_pos1h)]),
  fittedPOS = as.numeric(fitted(fitPOS1hd))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(df1hPOSd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "fittedPOS" = "darkgreen")) +
  ggtitle("ARIMA(0,1,1)")+
  theme_light()

grid.arrange(p1, p2, p3, ncol = 1)
```

### Auto-Regressive Neural Network (NNAR)

The NNAR (Neural Network AutoRegression) model uses lagged values of the series as inputs to a feed-forward neural network. It is a non-linear generalization of the AR model and can capture complex non-linear patterns in time series data.

#### Score Time Series

-   **Frequenza di 10 minuti**

```{r, fig.height=8}
nn10neg <- nnetar(score10neg)
nn10negDF <- data.frame(
  time = index(score10neg),
  neg = as.numeric(score10neg),
  NNAR = as.numeric(fitted(nn10neg))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(nn10negDF, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "NNAR" = "darkred")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

nn10neu <- nnetar(score10neu)
nn10neuDF <- data.frame(
  time = index(score10neu),
  neu = as.numeric(score10neu),
  NNAR = as.numeric(fitted(nn10neu))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(nn10neuDF, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "NNAR" = "darkblue")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

nn10pos <- nnetar(score10pos)
nn10posDF <- data.frame(
  time = index(score10pos),
  pos = as.numeric(score10pos),
  NNAR = as.numeric(fitted(nn10pos))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(nn10posDF, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "NNAR" = "darkgreen")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

grid.arrange(p1, p2, p3, ncol=1)
```

-   **Frequenza di 30 minuti**

```{r, fig.height=8}
nn30neg <- nnetar(score30neg)
nn30negDF <- data.frame(
  time = index(score30neg),
  neg = as.numeric(score30neg),
  NNAR = as.numeric(fitted(nn30neg))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(nn30negDF, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "NNAR" = "darkred")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

nn30neu <- nnetar(score30neu)
nn30neuDF <- data.frame(
  time = index(score30neu),
  neu = as.numeric(score30neu),
  NNAR = as.numeric(fitted(nn30neu))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(nn30neuDF, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "NNAR" = "darkblue")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

nn30pos <- nnetar(score30pos)
nn30posDF <- data.frame(
  time = index(score30pos),
  pos = as.numeric(score30pos),
  NNAR = as.numeric(fitted(nn30pos))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(nn30posDF, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "NNAR" = "darkgreen")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

grid.arrange(p1, p2, p3, ncol=1)
```

-   **Frequenza di 1 ora**

```{r, fig.height=8}
nn1hneg <- nnetar(score1hneg)
nn1hnegDF <- data.frame(
  time = index(score1hneg),
  neg = as.numeric(score1hneg),
  NNAR = as.numeric(fitted(nn1hneg))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(nn1hnegDF, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "NNAR" = "darkred")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

nn1hneu <- nnetar(score1hneu)
nn1hneuDF <- data.frame(
  time = index(score1hneu),
  neu = as.numeric(score1hneu),
  NNAR = as.numeric(fitted(nn1hneu))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(nn1hneuDF, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "NNAR" = "darkblue")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

nn1hpos <- nnetar(score1hpos)
nn1hposDF <- data.frame(
  time = index(score1hpos),
  pos = as.numeric(score1hpos),
  NNAR = as.numeric(fitted(nn1hpos))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(nn1hposDF, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "NNAR" = "darkgreen")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

grid.arrange(p1, p2, p3, ncol=1)
```

#### Temporal Derivative Time Series of the Score

-   **Frequenza di 10 minuti**

```{r, fig.height=8}
nn10negd <- nnetar(deriv_neg10)
nn10negDFd <- data.frame(
  time = index(deriv_neg10),
  neg = as.numeric(deriv_neg10),
  NNAR = as.numeric(fitted(nn10negd))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(nn10negDFd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "NNAR" = "darkred")) +
  ggtitle("NNAR")+
  theme_light()
# autoplot(forecast(nn10negd, PI=TRUE, h=10), include=100)

nn10neud <- nnetar(deriv_neu10)
nn10neuDFd <- data.frame(
  time = index(deriv_neu10),
  neu = as.numeric(deriv_neu10),
  NNAR = as.numeric(fitted(nn10neud))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(nn10neuDFd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "NNAR" = "darkblue")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

nn10posd <- nnetar(deriv_pos10)
nn10posDFd <- data.frame(
  time = index(deriv_pos10),
  pos = as.numeric(deriv_pos10),
  NNAR = as.numeric(fitted(nn10posd))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(nn10posDFd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "NNAR" = "darkgreen")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

grid.arrange(p1, p2, p3, ncol=1)
```

-   **Frequenza di 30 minuti**

```{r, fig.height=8}
nn30negd <- nnetar(deriv_neg30)
nn30negDFd <- data.frame(
  time = index(deriv_neg30),
  neg = as.numeric(deriv_neg30),
  NNAR = as.numeric(fitted(nn30negd))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(nn30negDFd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "NNAR" = "darkred")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

nn30neud <- nnetar(deriv_neu30)
nn30neuDFd <- data.frame(
  time = index(deriv_neu30),
  neu = as.numeric(deriv_neu30),
  NNAR = as.numeric(fitted(nn30neud))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(nn30neuDFd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "NNAR" = "darkblue")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

nn30posd <- nnetar(deriv_pos30)
nn30posDFd <- data.frame(
  time = index(deriv_pos30),
  pos = as.numeric(deriv_pos30),
  NNAR = as.numeric(fitted(nn30posd))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(nn30posDFd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "NNAR" = "darkgreen")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

grid.arrange(p1, p2, p3, ncol=1)
```

-   **Frequenza di 1 ora**

```{r, fig.height=8}
nn1hnegd <- nnetar(deriv_neg1h)
nn1hnegDFd <- data.frame(
  time = index(deriv_neg1h),
  neg = as.numeric(deriv_neg1h),
  NNAR = as.numeric(fitted(nn1hnegd))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p1<-ggplot(nn1hnegDFd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neg" = "grey70", "NNAR" = "darkred")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

nn1hneud <- nnetar(deriv_neu1h)
nn1hneuDFd <- data.frame(
  time = index(deriv_neu1h),
  neu = as.numeric(deriv_neu1h),
  NNAR = as.numeric(fitted(nn1hneud))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p2<-ggplot(nn1hneuDFd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("neu" = "grey70", "NNAR" = "darkblue")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

nn1hposd <- nnetar(deriv_pos1h)
nn1hposDFd <- data.frame(
  time = index(deriv_pos1h),
  pos = as.numeric(deriv_pos1h),
  NNAR = as.numeric(fitted(nn1hposd))
)%>%
  reshape2::melt(id.vars = "time", variable.name = "Series", value.name = "value")

p3<-ggplot(nn1hposDFd, aes(x = time, y = value, color = Series)) +
  geom_line() +
  scale_color_manual(values = c("pos" = "grey70", "NNAR" = "darkgreen")) +
  ggtitle("NNAR")+
  theme_light()
#autoplot(forecast(nn10neg, PI=TRUE, h=10))

grid.arrange(p1, p2, p3, ncol=1)
```

## Statistical Model Evaluation

To evaluate the models we adopt a hybrid Cross-Validation / Train-Test Split approach. This mixed strategy combines the advantages of **train-test split** and **time-series cross-validation** to obtain a more complete and reliable assessment of model performance. The temporal split into a **training set** and a **test set** simulates real-world conditions, evaluating the model's ability to generalise to unseen data with particular focus on long-horizon forecasts. Within the training set, **rolling-window cross-validation** (tsCV) tests the model on progressively expanding sub-windows, identifying potential overfitting issues and tuning parameters for improved short-term forecasts.

### Model Evaluation for Score Time Series

```{r, include=FALSE}
# List of score time series
score_list <- list(
  score10neg = score10neg,
  score10neu = score10neu,
  score10pos = score10pos,
  score30neg = score30neg,
  score30neu = score30neu,
  score30pos = score30pos,
  score1hneg = score1hneg,
  score1hneu = score1hneu,
  score1hpos = score1hpos
)

# Funzione per eseguire il train-test split
train_test_split <- function(series, test_size) {
  n <- length(series)
  train_end <- n - test_size
  train <- series[1:train_end]
  test <- series[(train_end + 1):n]
  return(list(train = train, test = test))
}


# Function to compute error metrics
calculate_metrics <- function(errors) {
  rmse <- sqrt(mean(errors^2, na.rm = TRUE))
  mae <- mean(abs(errors), na.rm = TRUE)
  return(tibble(RMSE = rmse, MAE = mae))
}

# Function to train a model and forecast on the test set
train_and_forecast <- function(train, test, model_type) {
  if (model_type == "ARIMA") {
    fit <- auto.arima(train, seasonal = FALSE, 
                      stepwise = FALSE, 
                      approximation = FALSE)
    forecasted <- forecast(fit, h = length(test))
  } else if (model_type == "NNETAR") {
    fit <- nnetar(train)
    forecasted <- forecast(fit, h = length(test))
  }
  errors <- test - forecasted$mean
  return(calculate_metrics(errors))
}

# Loop over all series and models
results_score <- list()
test_size <- 10  # Number of points reserved for the test set

for (series_name in names(score_list)) {
  series <- score_list[[series_name]]
  
  # Train-test split
  split <- train_test_split(series, test_size)
  train <- split$train
  test <- split$test
  
  # Forecast on test set with ARIMA
  arima_test_metrics <- train_and_forecast(train, test, "ARIMA")
  arima_test_metrics <- mutate(arima_test_metrics, Serie = series_name, Model = "ARIMA", Stage = "Test Set")
  

  # Forecast on test set with NNETAR
  nnetar_test_metrics <- train_and_forecast(train, test, "NNETAR")
  nnetar_test_metrics <- mutate(nnetar_test_metrics, Serie = series_name, Model = "NNETAR", Stage = "Test Set")
  
  # Append results
  results_score[[series_name]] <- bind_rows(arima_test_metrics,nnetar_test_metrics)
}

# Combine results into a single tibble
final_results_score <- bind_rows(results_score)

# Sort results by series, model, and stage
final_results_score <- final_results_score %>%
  arrange(Serie, Model, Stage)
```

We observe the results obtained by the models:

```{r}
# Print results
print(final_results_score)
```

We observe the best models on the **Test Set**:

```{r}
best_Score_models <- final_results_score %>%
  group_by(Serie) %>%
  slice_min(MAE, with_ties = FALSE) %>%
  ungroup()

print(best_Score_models)
```

We observe the forecast plots:

```{r, include=FALSE}
# Function to generate forecast data for plotting
generate_forecast_data <- function(series, test_size, model_type) {
  split <- train_test_split(series, test_size)
  train <- split$train
  test <- split$test
  
  if (model_type == "ARIMA") {
    fit <- auto.arima(train, seasonal = FALSE, 
                      stepwise = FALSE, 
                      approximation = FALSE)
    forecasted <- forecast(fit, h = length(test))
  } else if (model_type == "NNETAR") {
    fit <- nnetar(train)
    forecasted <- forecast(fit, h = length(test))
  }
  
  # Create a tibble with actual and forecast data
  forecast_data <- tibble(
    Time = seq_along(series),
    Value = c(train, test),
    Prediction = c(rep(NA, length(train)), forecasted$mean),
    Type = c(rep("Observed", length(train)), rep("Test", length(test)))
  )
  
  return(forecast_data)
}

# List to store data for plotting
plot_data_list <- list()

# Generate data for the series with best models
for (row in seq_len(nrow(best_Score_models))) {
  serie_name <- best_Score_models$Serie[row]
  model_type <- best_Score_models$Model[row]
  
  series <- score_list[[serie_name]]
  plot_data <- generate_forecast_data(series, test_size, model_type)
  plot_data <- mutate(plot_data, Serie = serie_name, Model = model_type)
  
  plot_data_list[[serie_name]] <- plot_data
}
```

```{r}
# Combine data into a single tibble
final_plot_data <- bind_rows(plot_data_list) %>%
  group_by(Serie) %>%
  filter(Time >= max(Time) - 40) %>%
  ungroup()

# Plot con ggplot2
ggplot(final_plot_data, aes(x = Time, y = Value)) +
  geom_line(aes(color = Type), size = 1.2) + 
  geom_line(aes(y = Prediction, linetype = "Forecast"), size = 1.1, color = "blue") +
  facet_wrap(~ Serie + Model, scales = "free") +
  labs(
    title = "Best Model Forecasts",
    x = "Tempo",
    y = "Valore",
    color = "Tipo",
    linetype = "Legenda"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12)
  )
```

### Model Evaluation for Derivative Time Series

```{r}
# List of derivative time series
deriv_list <- list(
  deriv10neg = deriv_neg10,
  deriv10neu = deriv_neu10,
  deriv10pos = deriv_pos10,
  deriv30neg = deriv_neg30,
  deriv30neu = deriv_neu30,
  deriv30pos = deriv_pos30,
  deriv1hneg = deriv_neg1h,
  deriv1hneu = deriv_neu1h,
  deriv1hpos = deriv_pos1h
)

# Loop over all series and models
results_deriv <- list()
test_size <- 10  # Number of points reserved for the test set

for (series_name in names(deriv_list)) {
  series <- deriv_list[[series_name]]
  
  # Train-test split
  split <- train_test_split(series, test_size)
  train <- split$train
  test <- split$test
  
  
  # Forecast on test set with ARIMA
  arima_test_metrics <- train_and_forecast(train, test, "ARIMA")
  arima_test_metrics <- mutate(arima_test_metrics, Serie = series_name, Model = "ARIMA", Stage = "Test Set")
  
  
  # Forecast on test set with NNETAR
  nnetar_test_metrics <- train_and_forecast(train, test, "NNETAR")
  nnetar_test_metrics <- mutate(nnetar_test_metrics, Serie = series_name, Model = "NNETAR", Stage = "Test Set")
  
  # Append results
  results_deriv[[series_name]] <- bind_rows(arima_test_metrics,nnetar_test_metrics)
}

# Combine results into a single tibble
final_results_deriv <- bind_rows(results_deriv)

# Sort results by series, model, and stage
final_results_deriv <- final_results_deriv %>%
  arrange(Serie, Model, Stage)
```

```{r}
print(final_results_deriv)
```

```{r}
best_deriv_models <- final_results_deriv %>%
  group_by(Serie) %>%
  slice_min(MAE, with_ties = FALSE) %>%
  ungroup()

print(best_deriv_models)
```

We observe the forecast plots::

```{r}
# List to store data for plotting
plot_data_list2 <- list()

# Generate data for the series with best models
for (row in seq_len(nrow(best_deriv_models))) {
  serie_name <- best_deriv_models$Serie[row]
  model_type <- best_deriv_models$Model[row]
  
  series <- deriv_list[[serie_name]]
  plot_data <- generate_forecast_data(series, test_size, model_type)
  plot_data <- mutate(plot_data, Serie = serie_name, Model = model_type)
  
  plot_data_list2[[serie_name]] <- plot_data
}

# Combine data into a single tibble
final_plot_data2 <- bind_rows(plot_data_list2) %>%
  group_by(Serie) %>%
  filter(Time >= max(Time) - 30) %>%
  ungroup()

# Plot con ggplot2
ggplot(final_plot_data2, aes(x = Time, y = Value)) +
  geom_line(aes(color = Type), size = 1.2) + 
  geom_line(aes(y = Prediction, linetype = "Forecast"), size = 1.1, color = "blue") +
  facet_wrap(~ Serie + Model, scales = "free") +
  labs(
    title = "Best Model Forecasts",
    x = "Tempo",
    y = "Valore",
    color = "Tipo",
    linetype = "Legenda"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12)
  )
```
